---
title: "EDT + Waze April 2017"
output:
  html_document:
    df_print: paged
---

## Working with linked EDT and Waze incident data for April 2017

Started 2017-12-07

Ideas:

- Frequency of events - histograms of different Waze event frequencies within a day, for different event types (accident, jam, weather hazard, road closed), and same for EDT incident count frequencies within a day
- Incident type by time of day
- Duration of incidents
- Spatial overlay of events - similar to Lia's slider idea, make a figure which builds up hotspots of Waze and EDT by day
- Spatiotemporal join - Paul's 2x2 for time (early, laggy) and space (near, far) matching of EDT and Waze, by type. In the future, using data snapped to HPMS, use roadway classification as well, or could use day of week.

```{r setup, message = F, warning = F, echo = F}
# Read in libraries and set working directory
knitr::opts_chunk$set(echo = F, warning=F, message=F)
options(width = 2400, stringsAsFactors = F)

library(tidyverse)
library(maps) # for mapping base layers
library(reshape)
library(DT) # for interatctive datatable
# devtools::install_github("ropensci/plotly") # for latest version
library(plotly) # do after ggplot2
library(sp)
library(maptools) # readShapePoly() and others
library(mapproj) # for coord_map()
library(rgdal) # for readOGR(), needed for reading in ArcM shapefiles
library(rgeos) # for gIntersection, to clip two shapefiles
library(knitr)
library(kableExtra) # formatting tables 

#Set file locations

if(length(grep("flynn", getwd())) > 0) {mappeddrive = "W:"} # change this to match the mapped network drive on your machine (Z: in original)
if(length(grep("sudderth", getwd())) > 0) {mappeddrive = "S:"} 

wazedir <- (file.path(mappeddrive,"SDI Pilot Projects/Waze"))
volpewazedir <- "//vntscex.local/DFS/Projects/PROJ-OR02A2/SDI/"

wazedir <- c("/Users/danflynn/Documents/Volpe Remote/tempdir/")
 
knitr::opts_knit$set(root.dir = wazedir)  # setwd(wazedir) # to develop code in working R session
```

```{r dataimport}
# Read in merged Waze and EDT data 
ew <- readRDS(file.path(wazedir, "merged.waze.edt.April_MD.rds"))
ww <- readRDS(file.path(wazedir, "merged.waze.waze.April_MD.rds"))


load(file.path(wazedir, "2017-04_1_CrashFact_edited.RData"))
# Waze: Comes from aggregated monthly Waze events, clipped to a 0.5 mile buffer around MD, for April 2017
load(file.path(wazedir, "MD_buffered__2017-04.RData"))

link <- read.csv(file.path(wazedir, "EDT_Waze_link_April_MD.csv"))
linkww <- read.csv(file.path(wazedir, "Waze_Waze_link_April_MD.csv"))


# Read in county data from Census
# counties <- readOGR(file.path(wazedir, "Working Documents/Census Files/Counties", layer = "."))
# md_counties <- counties[counties$STATEFP == 24,]

```

## Summary of EDT-Waze linkages


```{r matchtable}

## Calculations for EDT/Waze match table
Total.EDT.events <- length(unique(edt.april$ID))
EDT.events.with.Waze <- length(unique(link$id.edt))
EDT.events.without.Waze <- Total.EDT.events - EDT.events.with.Waze
Pct.EDT.match <- round(100*EDT.events.with.Waze/Total.EDT.events,2)

Total.Waze.events <- length(unique(d$uuid))
Waze.events.with.EDT <- length(unique(link$uuid.waze))
Waze.events.without.EDT <- Total.Waze.events - Waze.events.with.EDT
Pct.Waze.match <- round(100*Waze.events.with.EDT/Total.Waze.events,2)


mm <- matrix(c(EDT.events.with.Waze, EDT.events.without.Waze, Total.EDT.events, Pct.EDT.match,
         Waze.events.with.EDT, Waze.events.without.EDT,Total.Waze.events, Pct.Waze.match), byrow=T,
       nrow = 2, ncol = 4)

row.names(mm) <- c("EDT", "Waze")
colnames(mm) <- c("Matching", "Non-Matching", "Total", "Percent Matching")

kable(mm, format.args = list(big.mark = ",")) %>% kable_styling(bootstrap_options = c("striped"))
```

### EDT-Waze linkages by Urban/Rural

```{r matchbyurbanrural}
# Count of urbanrural

ew$UA <- !is.na(ew$EDT_UA_Name) | !is.na(ew$Waze_UA_Name)

mm <- aggregate(UA ~ match, 
                FUN = function(x) cbind(length(x), length(x[!x])),
                data = ew)

mm <- data.frame(mm[,1], as.data.frame(mm[,2]))

mm[,1] <- c("EDT only", "EDT-Waze Match", "Waze only")

mm$Pct.Urban <- round(100*mm[,2]/rowSums(mm[2:3]), 2)

colnames(mm) <- c("Match Type", "Urban", "Rural", "Percent Urban")

kable(mm, format.args = list(big.mark = ",")) %>% kable_styling(bootstrap_options = c("striped")) 


```

### EDT-Waze linkages by Urban/Rural and Waze Event Type

```{r matchbyurbanruraltype}
mm <- aggregate(UA ~ match + type, 
                FUN = function(x) cbind(length(x), length(x[!x])),
                data = ew)


mm <- data.frame(mm[,1:2], as.data.frame(mm[,3]))

mm$Pct.Urban <- mm[,3]/rowSums(mm[3:4])

mm[,1] <- rep(c("EDT-Waze Match", "Waze only"), length(unique(mm[,2])))

mm$Pct.Urban <- round(100*mm$Pct.Urban, 2)


colnames(mm) <- c("Match Type", "Waze Type", "Urban", "Rural", "Percent Urban")

# datatable(mm, filter = 'top') %>% 
#   formatCurrency(c('Urban', 'Rural'), currency = "", digits = 0) %>%
#   formatPercentage('Percent Urban', digits = 1)

kable(mm, format.args = list(big.mark = ",")) %>% kable_styling(bootstrap_options = c("striped")) 


```

## Summary of Waze-Waze linkages


```{r wwmatchtable}

## Calculations for EDT/Waze match table
Total.Waze.Incidents <- length(unique(d$uuid))
Accidents.with.all.Waze <- length(unique(linkww$id.accident))
Accidents.not.matching <- Total.Waze.Incidents - Accidents.with.all.Waze
Pct.match <- round(100*Accidents.not.matching/Total.Waze.Incidents,2)

Incidents.with.Accidents <- length(unique(linkww$id.incidents))
Pct.Waze.match.by.incident <- round(100*Incidents.with.Accidents/Total.Waze.Incidents,2)

mm <- matrix(c(Accidents.with.all.Waze, Total.Waze.Incidents, Pct.match,
         Incidents.with.Accidents, Total.Waze.Incidents, Pct.Waze.match.by.incident), byrow=T,
       nrow = 2, ncol = 3)

row.names(mm) <- c("Waze Accidents", "Waze Incidents")
colnames(mm) <- c("Matching", "Total", "Percent Matching")

kable(mm, format.args = list(big.mark = ",")) %>% kable_styling(bootstrap_options = c("striped"))
```

```{r wwmatchbyurbanrural, eval=F}
# Count of urbanrural

link.waze.waze$UA <- !is.na(ww$Waze_UA_Name) | !is.na(ww$Waze_UA_Name.y)

mm <- aggregate(UA ~ match, 
                FUN = function(x) cbind(length(x), length(x[!x])),
                data = link.waze.waze)

mm <- data.frame(mm[,1], as.data.frame(mm[,2]))

mm[,1] <- c("Waze-Waze Match", "Waze incident only")

mm$Pct.Urban <- round(100*mm[,2]/rowSums(mm[2:3]), 2)

colnames(mm) <- c("Match Type", "Urban", "Rural", "Percent Urban")

kable(mm, format.args = list(big.mark = ",")) %>% kable_styling(bootstrap_options = c("striped")) 


```

```{r wwmatchbyurbanruraltype, eval=F}
mm <- aggregate(UA ~ match + type, 
                FUN = function(x) cbind(length(x), length(x[!x])),
                data = link.waze.waze)


mm <- data.frame(mm[,1:2], as.data.frame(mm[,3]))

mm$Pct.Urban <- mm[,3]/rowSums(mm[3:4])

mm[,1] <- rep(c("EDT-Waze Match", "Waze only"), length(unique(mm[,2])))

colnames(mm) <- c("Match Type", "Waze Type", "Urban", "Rural", "Percent Urban")

datatable(mm, filter = 'top') %>% 
  formatCurrency(c('Urban', 'Rural'), currency = "", digits = 0) %>%
  formatPercentage('Percent Urban', digits = 1)
```
