---
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
      smooth_scroll: yes
author: "Dan Flynn | Volpe"
title: "Waze / Economic Shock Test"
---


<!-- saved from url=(0014)about:internet -->
<!-- This needs to be manually added to the second line of the html -->


```{r setup, echo = F, message = F, warning=F}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = F)
library(tidyverse)
library(DT)
library(lubridate)
library(sp)
library(maps) # for mapping base layers
library(rgdal) # for readOGR(), needed for reading in ArcM shapefiles
library(rgeos) # for gIntersection, to clip two shapefiles
library(kableExtra)

user <- paste0( "/home/", system("whoami", intern = TRUE))
localdir <- paste0(user, "/workingdata/") # full path for readOGR
teambucket <- "prod-sdc-sdi-911061262852-us-east-1-bucket"
output.loc <- "~/tempout"
codeloc <- "~/SDI_Waze"

```

### Overview

The goal of this quick analysis is the following:

- Visualize Waze report data for example counties impacted by one natural disaster, in this case Hurricane Florence
- Compare counts of total Waze incident reports for different alert types by day, by county, for 2018 and 2017

These visuals will help assess what could be drawn from the Waze alert table for to support economic activity models for the Board of Governors of the Federal Reserve System. Note: in addition to the alert table, additional system-generated data on jams, including spatial extent of jam and speed, are also available. These visuals only focus on the alert table, which includes user reports of crashes, weather hazards, hazards on the road, road closures, and user-reported jams.

In brief, this involved querying the Waze alert data for North Carolina, and they using GIS to assign with county FIPS codes to each alert. The data were then summarized as counts of Waze reports, by alert types, day, and county. The data were then further filtered to selected counties with varying levels of impact to Hurricane Florence, using [FEMA disaster declaration](https://www.census.gov/topics/preparedness/events/hurricanes/florence.html). Twenty-nine counties in North Carolina were included in the disaster declaration out of the 100 counties in the state. Six states and the District of Columbia were potentially affected. Disaster was declared on September 14, 2018, and the official incident period was September 7 to 29, 2018. No disasters were declared in North Carolina in 2017.

```{r prep_data}
state = "NC"

# Using output from in ReduceWaze_SDC.R
months_2017 = 5:11

if(length(grep("NC_Econ_Shock.RData", dir(file.path(localdir, state))))==0){
  r2017 = vector()
  for(i in months_2017){
    load(file.path(localdir, state, paste0("NC_2017-", formatC(i, width = 2, flag = 0),".RData")))
    r2017 = rbind(r2017, mb)
    }
  
  months_2018 = 5:11
  
  r2018 = vector()
  for(i in months_2018){
    load(file.path(localdir, state, paste0("NC_2018-", formatC(i, width = 2, flag = 0),".RData")))
    r2018 = rbind(r2018, mb)
    }
    
  # Select columns to keep
  keep = c("alert_uuid", "alert_type", "sub_type", "lon", "lat",
           "nrecord",
           "city", "time","last.pull.time", "road_type")
  
  d = rbind(r2017[keep],
            r2018[keep])
  
  rm(mb, r2017, r2018)
  
  # Get spatial data
  counties <- rgdal::readOGR(file.path(localdir, "census"), layer = "cb_2017_us_county_500k")
  
  
  # Project to Albers equal area, ESRI 102008
  proj <- showP4(showWKT("+init=epsg:102008"))
  # USGS version of AEA. Use this for all projections for consistency; this is what the hexagon layer is in 
  proj.USGS <- "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0"
  
  counties <- spTransform(counties, CRS(proj.USGS))
  
  FIPS = maps::state.fips %>% filter(abb %in% state)
  FIPS = FIPS[!duplicated(FIPS$fips),c("abb", "fips")]
  names(FIPS) = c("state", "FIPS")
  FIPS$FIPS <- formatC(FIPS$FIPS, width = 2, flag = "0")
  
  
  # County overlay ----
  # d now has all NC events for selected months. 6 M rows, 10 columns 
  # want to get the county for each event
  d_s <- SpatialPointsDataFrame(coords = cbind(d["lon"], d["lat"]),
                                data = d,
                                proj4string = CRS("+proj=longlat +datum=WGS84"))
  
  d_s <- spTransform(d_s, CRS(proj.USGS))
  
  co_intersect <- gIntersects(d_s, counties, byid = T, returnDense = F)
  
  cv <- unlist(co_intersect)
  
  cf <- as.character(counties@data$GEOID[cv])
  names(cf) = names(cv)
  
  d_s$county_FIPS <- cf[match(rownames(d_s@data), names(cf))]
  
  d_s <- d_s[!is.na(d_s$county_FIPS),]
  
  save(d_s, file = file.path(localdir, state, "NC_Econ_Shock.RData"))
  } else {
  load(file.path(localdir, state, "NC_Econ_Shock.RData"))
}

```

Selected counties:

```{r select_counties}
# Which counties do we want to focus on?

co_name = c("Brunswick",
            "Carteret",
            "Hyde",
            "Moore",
            "Johnston",
            "Robeson",
            "Buncombe",
            "Rockingham",
            "Haywood")

co_type = c(rep("Outer Banks - Disaster", 3),
            rep("Inland - Disaster", 3),
            rep("Inland - No Disaster", 3))

co_fips = c(37019,
            37031,
            37095,
            37125,
            37101,
            37155,
            37021,
            37157,
            37087)

co <- data.frame(co_name, co_type, co_fips)


ds <- d_s@data[d_s@data$county_FIPS %in% co$co_fips,]
ds$Day <- format(ds$time, "%Y-%m-%d")
ds$Year <- format(ds$time, "%Y")
ds$JDay <- format(ds$time, "%j")

# Summarize to counts of alert_types by day
dz <- ds %>%
  group_by(county_FIPS, Year, Day, JDay, alert_type) %>%
  summarize(count = n())

co$co_fips = as.character(co$co_fips)

dz2 <- left_join(dz, co, by = c("county_FIPS" = "co_fips"))

dz3 <- dz2 %>%
  group_by(co_name, co_type, county_FIPS) %>%
  summarize(count = sum(count))
datatable(dz3, caption = "Selected counties in North Carolina, with total counts of Waze incident reports in May-November of 2017 and 2018",
          rownames = F,
          options = list(dom = "t",
                         order = list(list(1, 'desc')))) %>%
          formatCurrency(4, currency="", digits = 0)  
          
```

Using data from May to November of each 2017 and 2018, in total there were `r format(nrow(d_s), big.mark = ",")` Waze reports in NC. 

The vertical line in the plots below mark the disaster declaration on September 14, 2018. Smoothed lines are splines.

### Outer Banks, Disaster declared

```{r plot1}

# ggplot(dz %>% filter(Year == "2017")) +
#   geom_point(aes(x = JDay, y = count, group = alert_type, color = alert_type)) +
#   facet_wrap(~county_FIPS)
# 
# ggplot(dz %>% filter(Year == "2018")) +
#   geom_point(aes(x = JDay, y = count, group = alert_type, color = alert_type)) +
#   facet_wrap(~county_FIPS)

class(dz$JDay) = "numeric"
vline.dat <- data.frame(Year = unique(dz$Year), vl = c(NA, 257))
spanit = 0.5 # default = 1, smaller values = wigglier

# Brunswick 
ggplot(dz %>% filter(county_FIPS == "37019"),
       aes(x = JDay, y = count, group = alert_type, color = alert_type)) +
  geom_point(alpha = 0.2) + geom_smooth(span=spanit) +
  #ylim(c(0, 250))+
  facet_wrap(~Year) +
  ggtitle("Brunswick County - Outer Banks, Disaster declared") + 
  xlab("Day of year") +
  geom_vline(aes(xintercept = vl), data = vline.dat)

# ggplot(dz %>% filter(county_FIPS == "37019" & alert_type == "ACCIDENT"),
#        aes(x = JDay, y = count, group = alert_type, color = alert_type)) +
#   geom_point(alpha = 0.2) + geom_smooth(span = spanit) +
#   #ylim(c(0, 250))+
#   facet_wrap(~Year) +
#   ggtitle("Brunswick County - Outer Banks, Disaster declared") + 
#   xlab("Day of year") +
#   geom_vline(aes(xintercept = vl), data = vline.dat)

# Carteret
ggplot(dz %>% filter(county_FIPS == "37031"),
       aes(x = JDay, y = count, group = alert_type, color = alert_type)) +
  geom_point(alpha = 0.2) + geom_smooth(span=spanit) +
  facet_wrap(~Year) +
  ggtitle("Carteret County - Outer Banks, Disaster declared") + 
  xlab("Day of year") +
  geom_vline(aes(xintercept = vl), data = vline.dat)

# 
# ggplot(dz %>% filter(county_FIPS == "37031" & alert_type == "ACCIDENT"),
#        aes(x = JDay, y = count, group = alert_type, color = alert_type)) +
#   geom_point(alpha = 0.2) + geom_smooth(span=spanit) +
#   facet_wrap(~Year) +
#   ggtitle("Carteret County - Outer Banks, Disaster declared") + 
#   xlab("Day of year") +
#   geom_vline(aes(xintercept = vl), data = vline.dat)

# Hyde
ggplot(dz %>% filter(county_FIPS == "37095"),
       aes(x = JDay, y = count, group = alert_type, color = alert_type)) +
  geom_point(alpha = 0.2) + geom_smooth(span=0.5) +
  ylim(c(0, 50))+
  facet_wrap(~Year) +
  ggtitle("Hyde County - Outer Banks, Disaster declared")  +
  xlab("Day of year") +
  geom_vline(aes(xintercept = vl), data = vline.dat)
```

### Inland, Disaster declared

```{r plot2}
ggplot(dz %>% filter(county_FIPS == "37101"),
       aes(x = JDay, y = count, group = alert_type, color = alert_type)) +
  geom_point(alpha = 0.2) + geom_smooth(span=spanit) +
#  ylim(c(0, 150))+
  facet_wrap(~Year) +
  ggtitle("Johnston County - Inland, Disaster declared") + 
  xlab("Day of year") +
  geom_vline(aes(xintercept = vl), data = vline.dat)

ggplot(dz %>% filter(county_FIPS == "37125"),
       aes(x = JDay, y = count, group = alert_type, color = alert_type)) +
  geom_point(alpha = 0.2) + geom_smooth(span=spanit) +
  ylim(c(0, 150))+
  facet_wrap(~Year) +
  ggtitle("Moore County - Inland, Disaster declared") + 
  xlab("Day of year") +
  geom_vline(aes(xintercept = vl), data = vline.dat)

ggplot(dz %>% filter(county_FIPS == "37155"),
       aes(x = JDay, y = count, group = alert_type, color = alert_type)) +
  geom_point(alpha = 0.2) + geom_smooth(span=spanit) +
#  ylim(c(0, 150))+
  facet_wrap(~Year) +
  ggtitle("Robeson County - Inland, Disaster declared") + 
  xlab("Day of year") +
  geom_vline(aes(xintercept = vl), data = vline.dat)



```



### Inland, No disaster declared

```{r plot3}
# Buncombe (Asheville)
ggplot(dz %>% filter(county_FIPS == "37021"),
       aes(x = JDay, y = count, group = alert_type, color = alert_type)) +
  geom_point(alpha = 0.2) + geom_smooth(span=spanit) +
  #ylim(c(0, 250))+
  facet_wrap(~Year) + 
  ggtitle("Buncombe County (Asheville) - Less Affected") + 
  xlab("Day of year") +
  geom_vline(aes(xintercept = vl), data = vline.dat)


# Haywood
ggplot(dz %>% filter(county_FIPS == "37087"),
       aes(x = JDay, y = count, group = alert_type, color = alert_type)) +
  geom_point(alpha = 0.2) + geom_smooth(span=spanit) +
  #ylim(c(0, 250))+
  facet_wrap(~Year) + 
  ggtitle("Haywood County - Less Affected") + 
  xlab("Day of year") +
  geom_vline(aes(xintercept = vl), data = vline.dat)

# Rockingham
ggplot(dz %>% filter(county_FIPS == "37157"),
       aes(x = JDay, y = count, group = alert_type, color = alert_type)) +
  geom_point(alpha = 0.2) + geom_smooth(span=spanit) +
  #ylim(c(0, 250))+
  facet_wrap(~Year) + 
  ggtitle("Rockingham County - Less Affected") + 
  xlab("Day of year") +
  geom_vline(aes(xintercept = vl), data = vline.dat)


```

