---
title: "EDT + Waze Joining: August Example"
output:
  html_notebook:
    df_print: paged
  html_document:
    df_print: paged
---

Initial work in mapping EDT and Waze incident data, using week of August 1 as example

Started 2017-11-28

Ideas:

- Frequency of events - histograms of different Waze event frequencies within a day, for different event types (accident, jam, weather hazard, road closed), and same for EDT incident count frequencies within a day
- Incident type by time of day
- Duration of incidents
- Spatial overlay of events - similar to Lia's slider idea, make a figure which builds up hotspots of Waze and EDT by day
- Spatiotemporal join - Paul's 2x2 for time (early, laggy) and space (near, far) matching of EDT and Waze, by type. In the future, using data snapped to HPMS, use roadway classification as well, or could use day of week.

Summarizing the overlay. How many:
- Waze events match EDT events within the specified buffers? (M)
- Waze events within the state do not match EDT events? (W)
- EDT events within the state do not match Waze events? (E)

FOr each of these three categories, what are the proportions for:
- Different roadway types
- Rural/urban areas
- Types of events (severe/fatal in EDT, accident vs non-accient for Waze)

This document usese the August data as an example, since these day have been worked with before and use the same coordinate system and projection.

```{r setup, message = F, warning = F}
# Read in libraries and set working directory
knitr::opts_chunk$set(echo = T, warning=F, message=F)
options(width = 2400, stringsAsFactors = F)

library(tidyverse)
library(maps) # for mapping base layers
library(DT) # for datatable
# devtools::install_github("ropensci/plotly") # for latest version
library(plotly) # do after ggplot2
library(sp)
library(maptools) # readShapePoly() and others
library(mapproj) # for coord_map()
library(rgdal) # for readOGR(), needed for reading in ArcM shapefiles
library(rgeos) # for gIntersection, to clip two shapefiles
library(raster)

wazedir <- "W:/SDI Pilot Projects/Waze/Working Documents"
knitr::opts_knit$set(root.dir = wazedir) 
```

```{r dataimport}
# Initial run: Start with the projected EDT and Waze data that Lia has prepared. It appears this is just the one week of August data for MD; use as a pilot, expand to other periods after (month of April). Clean these data and save as .RData. If already done, load the processed data.

if(length(grep('EDT_Waze.RData', dir("R/"))) < 1){
  waz.shp <- readOGR(dsn = ".", "Waze_sample_projected")
  edt.shp <- readOGR(dsn = ".", "EDT_sample_projected")
  
  # Fix time values: convert pub_millis to usable POSIX date class
  waz.shp$time <- as.POSIXct(waz.shp$pub_millis/1000, origin = "1970-01-01", tz="America/New_York")
  # Convert crashdate and time in EDT to POSIX
  edt.time <- paste(as.character(edt.shp$CrashDate), " ",
                    edt.shp$HourofDay, ":",
                    formatC(as.numeric(as.character(edt.shp$MinuteofDa)), width = 2, flag = "0"), sep = "")
  edt.shp$time <- strptime(edt.time, format = "%Y/%m/%d %H:%M", tz="America/New_York")

  # Read in county data from Census
  counties <- readOGR("CensusCounty/.")
  md_counties <- counties[counties$STATEFP == 24,]

  # Save imported waze and EDT files as .RData for faster import after the first time
  save(file = "R/EDT_Waze.RData", list = c("waz.shp", "edt.shp", "md_counties"))
} else { load('R/EDT_Waze.RData') }


```

### Initial mapping work

```{r fort.map, eval=FALSE, include=FALSE}
# Currently not run -- this is the start of making a fortified dataframe for ggplot mapping, to turn into a plotly map.
# Get points in polygons: events in each county
proj4string(edt.shp) <- proj4string(waz.shp) <- proj4string(md_counties)

md_pip <- over(waz.shp, md_counties[,"NAME"])

waz.shp$NAME <- md_pip

md_base <- ggplot(md_counties) + geom_map(data = md_counties, map = md_counties,
                                            aes(x = long, y = lat, group = group, map_id = id),
                                            color = "black", fill = NA) +
  ylab(label="") + xlab(label="") +
  theme_minimal() +
  coord_map(projection="albers", lat = 39, lat1 = 45)
```

```{r plotlymap}
counties <- map_data("county", "maryland")

waz.shp$tooltiptext <- with(waz.shp@data, paste(city, alert_type, time, sep = "\n"))
waz.shp$tooltiptext <- sub("NA", "", waz.shp$tooltiptext)

edt.shp$tooltiptext <- with(edt.shp@data, paste(County, HighestInj, time, sep = "\n"))
edt.shp$tooltiptext <- sub("NA", "", edt.shp$tooltiptext)


map.p <- ggplot() +
  ggtitle("2017-08-01 Spatial Overlay") +
  geom_polygon(data = counties, aes(x = long, y = lat, group = group), 
               color = "white", fill = "grey80") + theme_void() +
coord_fixed(1.3) +

  geom_point(data = waz.shp[format(waz.shp$time, "%d") == "01",]@data, # 2017-08-01 first
             aes(
               x = longitude, 
               y = latitude, 
               color = alert_type,
                text = tooltiptext),
             shape = 16) +

  geom_point(data = edt.shp[format(edt.shp$time, "%d") == "01",]@data, # 2017-08-01 first
             aes(
               x = GPSLong_Ne, 
               y = GPSLat, 
               text = tooltiptext,
               color = 'EDT'),
             shape = 16) +
  scale_color_brewer(palette = "Set1",
                        name = "Waze Alert Type and EDT Crash")

  
ggplotly(map.p, tooltip = "text", width = 1000)

```

### Interactive table of Waze events by day 

```{r freqevents}
# make a table of frequency of alert_type by day
freq.tab <- aggregate(uuid ~ format(time, "%d") + alert_type,
          FUN = length,
          data = waz.shp@data)

datatable(freq.tab,
          filter = 'top',
          colnames = c("Day" = 1, "Alert Type" = 2, "Count"= 3),
          rownames = F,
          options = list(dom = "ftip"))
```


### Spatiotemporal join

```{r stjoin, warning = F, eval = T} 

# for each EDT event, find the number and distance for nearest neighbors in space and time from Waze data
ex <- edt.shp@data

distbreaks <- c(0, 0.1, 0.5, 1, 5, 10, 15, 25, 50, 100, 300, 1000)
distcounts <- vector()

for(i in 1:nrow(ex)){

  daterangemin <- ex[i, 'time'] - 2*60*60
  daterangemax <- ex[i, 'time'] + 2*60*60
    
  # find waze events within 3 hrs of this crash
  
  wx <- waz.shp[waz.shp$time > daterangemin & waz.shp$time < daterangemax,]
  
  if(nrow(wx) ==0) {
    b <- rep(0, length(distbreaks))
  } else {
  w_sp <- SpatialPoints(coords = data.frame(wx$longitude, wx$latitude))
  e_sp <- SpatialPoints(coords = data.frame(ex[i,'GPSLong_Ne'], ex[i,'GPSLat']))
  
  t1 <- spDists(w_sp, e_sp, longlat = T) # distance in km to each edt event
  
  # Sum of events with 0.1 to 100 mi in various chunks
  b <- hist(t1, breaks = distbreaks, plot = F)$counts
  }

  distcounts <- rbind(distcounts, b)
  
  # Sum of events within 1 to 120 minutes
  
  }

colnames(distcounts) = distbreaks[2:length(distbreaks)]
rownames(distcounts) = ex$ID

dm <- melt(distcounts)
colnames(dm) <- c("ID", "Bin", "Count")
dm$Bin <- as.factor(dm$Bin)

gp <- ggplot(dm) + geom_histogram(aes(x = Count), bins = 50) + facet_wrap(~Bin) + ylab("Frequency") + xlab("Count of Waze Events")

ggplotly(gp, width = 1000, tooltip = c("count"))

```