---
title: ''
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

codeloc <- "~/SDI_Waze" 
teambucket <- "s3://prod-sdc-sdi-911061262852-us-east-1-bucket"
user <- paste0( "/home/", system("whoami", intern = TRUE)) # the user directory to use
localdir <- paste0(user, "/workingdata") # full path for readOGR
outputdir <- file.path(localdir, 'Random_Forest_Output')

library(tidyverse)
library(knitr)
library(kableExtra)
library(randomForest)
library(pander)
library(DT)
# Run this if you don't have these packages:
# source(file.path(codeloc, 'utility/get_packages.R'))

knitr::opts_knit$set(root.dir = localdir)

# read utility functions
source(file.path(codeloc, 'utility/wazefunctions.R'))

# read random forest functions
source(file.path(codeloc, "analysis/RandomForest_WazeGrid_Fx.R"))
```

```{r getfiles, message=FALSE, warning=FALSE}

# Pull down model outputs from S3 if necessary
md.files <-  c(
  'MD_VMT_Output_to_30.RData',
  'MD_Model_30_RandomForest_Output.RData',
  'MD_Model_63_RandomForest_Output.RData',
  'MD_Model_62_RandomForest_Output.RData',
  'MD_Model_61_RandomForest_Output.RData')

# MD files
#  system(paste("aws s3 ls", file.path(teambucket, 'MD/')))

for(i in md.files){
  if(length(grep(i, dir(outputdir))) == 0){
    system(paste("aws s3 cp",
                 file.path(teambucket, "MD", i),
                 file.path(outputdir, i)))
  }
}

ct.files <-  c(
  'CT_VMT_Output_to_30.RData',
  'CT_Model_30_RandomForest_Output.RData',
  'CT_Model_63_RandomForest_Output.RData',
  'CT_Model_62_RandomForest_Output.RData',
  'CT_Model_61_RandomForest_Output.RData')

# CT files
#  system(paste("aws s3 ls", file.path(teambucket, 'CT/')))

for(i in ct.files){
  if(length(grep(i, dir(outputdir))) == 0){
    system(paste("aws s3 cp",
                 file.path(teambucket, "CT", i),
                 file.path(outputdir, i)))
  }
}

```

# VMT versus AADT


- Average annual daily traffic (AADT), by sum of the AADT in the roads in a grid cell, from HPMS.

- VMT: Volume of traffic distributed by hour of day, day of week, and month of year from ATR data. Approximately 5,000 stations on the National Highway System (NHS). Hour of year VMT is the product of hour of year volume and HPMS segment length.

#### Maryland VMT model comparisons

```{r MD_VMT_metrics, echo = F}
# Get outputs
load(file.path(outputdir, "MD_VMT_Output_to_30.RData"))

tabl <- vector()
for(i in c("18", "61", "62", "30")){ tabl <- rbind(tabl, c(100*keyoutputs[[i]]$diag, round(keyoutputs[[i]]$auc, 4)))}

colnames(tabl)[1:5] = c("Accuracy", "Precision", "Recall", "False Positive Rate", "AUC")
rownames(tabl) =  c("18 - Simple", "61 - AADT", "62 - VMT", "30 - Full with VMT")
datatable(tabl, 
          caption = "Maryland VMT model comparisons",
          rownames = T,
          options = list(dom = "t",
                         order = list(list(5, 'desc')),
                         pageLength = 5)
          ) %>% formatCurrency(1:4, currency = "", digits = 2) %>% formatCurrency(5, currency = "", digits = 3) %>%
          formatStyle(1, background = styleEqual(max(tabl[,1]), 'lightgreen'))%>%
          formatStyle(2, background = styleEqual(max(tabl[,2]), 'lightgreen'))%>%
          formatStyle(3, background = styleEqual(max(tabl[,3]), 'lightgreen'))%>%
          formatStyle(4, background = styleEqual(min(tabl[,4]), 'lightgreen'))%>%
          formatStyle(5, background = styleEqual(max(tabl[,5]), 'lightgreen'))
```   

#### Connecticut VMT model comparisons


```{r CT_VMT_metrics, echo = F}
# Get outputs

load(file.path(localdir, "CT_VMT_Output_to_18"))
k_18 = keyoutputs
load(file.path(outputdir, "CT_VMT_Output_to_30.RData"))
keyoutputs[['18']] = k_18[['18']]

tabl <- vector()
for(i in c("18","61", "62",  "30")){ tabl <- rbind(tabl, c(100*keyoutputs[[i]]$diag, round(keyoutputs[[i]]$auc, 4)))}

colnames(tabl)[1:5] = c("Accuracy", "Precision", "Recall", "False Positive Rate", "AUC")
rownames(tabl) =  c("18 - Simple", "61 - AADT", "62 - VMT",  "30 - Full with VMT")
datatable(tabl, 
          caption = "Connecticut VMT model comparisons",
          rownames = T,
          options = list(dom = "t",
                         order = list(list(5, 'desc')),
                         pageLength = 5)
          ) %>% formatCurrency(1:4, currency = "", digits = 2) %>% 
                formatCurrency(5, currency = "", digits = 3) %>%
          formatStyle(1, background = styleEqual(max(tabl[,1]), 'lightgreen'))%>%
          formatStyle(2, background = styleEqual(max(tabl[,2]), 'lightgreen'))%>%
          formatStyle(3, background = styleEqual(max(tabl[,3]), 'lightgreen'))%>%
          formatStyle(4, background = styleEqual(min(tabl[,4]), 'lightgreen'))%>%
          formatStyle(5, background = styleEqual(max(tabl[,5]), 'lightgreen'))
```   


#### Utah VMT model comparisons

```{r UT_VMT_metrics, echo = F}
# Get outputs
load(file.path(localdir, "UT_VMT_Output_to_30.RData"))

tabl <- vector()
for(i in c("18","61", "62", "30")){ tabl <- rbind(tabl, c(100*keyoutputs[[i]]$diag, round(keyoutputs[[i]]$auc, 4)))}

colnames(tabl)[1:5] = c("Accuracy", "Precision", "Recall", "False Positive Rate", "AUC")
rownames(tabl) =  c("18 - Simple", "61 - AADT", "62 - VMT",  "30 - Full with VMT")
datatable(tabl, 
          caption = "Utah VMT model comparisons",
          rownames = T,
          options = list(dom = "t",
                         order = list(list(5, 'desc')),
                         pageLength = 5)
          ) %>% formatCurrency(1:4, currency = "", digits = 2) %>% 
                formatCurrency(5, currency = "", digits = 3) %>%
          formatStyle(1, background = styleEqual(max(tabl[,1]), 'lightgreen'))%>%
          formatStyle(2, background = styleEqual(max(tabl[,2]), 'lightgreen'))%>%
          formatStyle(3, background = styleEqual(max(tabl[,3]), 'lightgreen'))%>%
          formatStyle(4, background = styleEqual(min(tabl[,4]), 'lightgreen'))%>%
          formatStyle(5, background = styleEqual(max(tabl[,5]), 'lightgreen'))
```   


#### Virginia VMT model comparisons

```{r VA_VMT_metrics, echo = F}
# Get outputs
load(file.path(localdir, "VA_VMT_Output_to_30.RData"))

tabl <- vector()
for(i in c("18","61", "62", "30")){ tabl <- rbind(tabl, c(100*keyoutputs[[i]]$diag, round(keyoutputs[[i]]$auc, 4)))}

colnames(tabl)[1:5] = c("Accuracy", "Precision", "Recall", "False Positive Rate", "AUC")
rownames(tabl) =  c("18 - Simple", "61 - AADT", "62 - VMT",  "30 - Full with VMT")
datatable(tabl, 
          caption = "Virginia VMT model comparisons",
          rownames = T,
          options = list(dom = "t",
                         order = list(list(5, 'desc')),
                         pageLength = 5)
          ) %>% formatCurrency(1:4, currency = "", digits = 2) %>% 
                formatCurrency(5, currency = "", digits = 3) %>%
          formatStyle(1, background = styleEqual(max(tabl[,1]), 'lightgreen'))%>%
          formatStyle(2, background = styleEqual(max(tabl[,2]), 'lightgreen'))%>%
          formatStyle(3, background = styleEqual(max(tabl[,3]), 'lightgreen'))%>%
          formatStyle(4, background = styleEqual(min(tabl[,4]), 'lightgreen'))%>%
          formatStyle(5, background = styleEqual(max(tabl[,5]), 'lightgreen'))
```   



## Importance of VMT variables



```{r CT_MD_varimp_table, eval=T}
load(file.path(outputdir, "CT_Model_30_RandomForest_Output.RData"))
CT_rf.out = rf.out

load(file.path(outputdir, "MD_Model_30_RandomForest_Output.RData"))
MD_rf.out = rf.out

load(file.path(outputdir, "UT_Model_30_RandomForest_Output.RData"))
UT_rf.out = rf.out

load(file.path(outputdir, "VA_Model_30_RandomForest_Output.RData"))
VA_rf.out = rf.out

nimp = 100

CT_rf.out$importance = 100*CT_rf.out$importance / sum(CT_rf.out$importance)
MD_rf.out$importance = 100*MD_rf.out$importance / sum(MD_rf.out$importance)
UT_rf.out$importance = 100*UT_rf.out$importance / sum(UT_rf.out$importance)
VA_rf.out$importance = 100*VA_rf.out$importance / sum(VA_rf.out$importance)

CT_imp = CT_rf.out$importance[order(CT_rf.out$importance, decreasing = T)[1:nimp],]
MD_imp = MD_rf.out$importance[order(MD_rf.out$importance, decreasing = T)[1:nimp],]
UT_imp = UT_rf.out$importance[order(UT_rf.out$importance, decreasing = T)[1:nimp],]
VA_imp = VA_rf.out$importance[order(VA_rf.out$importance, decreasing = T)[1:nimp],]

CT_imp = data.frame(var = names(CT_imp),
                    CT_rank = 1:nimp,
                    CT_Importance = CT_imp)

MD_imp = data.frame(var = names(MD_imp),
                    MD_rank = 1:nimp,
                    MD_Importance = MD_imp)

UT_imp = data.frame(var = names(UT_imp),
                    UT_rank = 1:nimp,
                    UT_Importance = UT_imp)
VA_imp = data.frame(var = names(VA_imp),
                    VA_rank = 1:nimp,
                    VA_Importance = VA_imp)


CT_MD_imp = full_join(CT_imp, MD_imp, by = "var")
All_imp = full_join(CT_MD_imp, UT_imp, by = "var")
All_imp = full_join(All_imp, VA_imp, by = "var")

# Get rank of VMT variables
VMT_table = All_imp[grep("HOURLY", All_imp$var),]

VMT_table = VMT_table[order(VMT_table$var),]
VMT_table$var = c("Hourly VMT - 1 Interstate",
                  "Hourly VMT - 2 Principal Arterial - Freeway and Expressway",
                  "Hourly VMT - 3 Principal Arterial",
                  "Hourly VMT - 4 Minor Arterial",
                  "Hourly VMT - 5 Major Collector"
                  )

datatable(VMT_table,
          rownames = FALSE,
          colnames = c("VMT Variable", "CT Rank", "CT Importance",
                       "MD Rank", "MD Importance",
                       "UT Rank", "UT Importance",
                       "VA Rank", "VA Importance"),
          caption = "Importance of VMT variables for the four focal states.",
           options = list(dom = "t",
                          order = list(0, 'asc'),
                          autoWidth = T)
          ) %>% formatCurrency(c(3, 5, 7, 9), currency = "", digits = 2)

```


### Estimated crashes with and without VMT variables



