---
title: "Waze_coverage"
output: html_document
---

```{r setup, echo = F, message=FALSE}
knitr::opts_chunk$set(message = FALSE)
library(tidyverse) 
library(lubridate)
library(DT)

# Location of SDC SDI Waze team S3 bucket. Files will first be written to a temporary directory in this EC2 instance, then copied to the team bucket.
teambucket <- "prod-sdc-sdi-911061262852-us-east-1-bucket"
output.loc <- "~/tempout"
codeloc <- "~/SDI_Waze"
```

# Overview of Waze data coverage for all states

Can also look at EDT states
Will have to clip to borders for each state
Possibly multiple tables for large states

### Metrics
- Date of first and last entry in alert table for each state
- pct of 5 min intervals complete for each state, for time range April 1 2017 - March 30 2018

### Outputs

- number of major/minor accidents per day by state, by county
- Saved as flat files, to be sent to ATA for Tableau visualization

## Connect to Redshift and build test query
```{r}
source(file.path(codeloc, 'utility/connect_redshift_pgsql.R'))
```

### Test states in Redshift
```{r statealert_n, cache=TRUE}
# How many records are there in each state, total for last year. ~ 55 min

state_alert_query <- paste0("SELECT COUNT(DISTINCT alert_uuid), state, alert_type
                            FROM alert 
WHERE pub_utc_timestamp BETWEEN to_timestamp('2017-04-01 00:00:00','YYYY-MM-DD HH24:MI:SS') AND to_timestamp('2018-03-30 23:59:59','YYYY-MM-DD HH24:MI:SS')
                            GROUP BY state, alert_type;")
system.time(results <- dbGetQuery(conn, state_alert_query))

write.csv(results, "~/workingdata/State_Alert-Type_Count.csv", row.names= F)
```

```{r DTresults1}
datatable(results, filter = 'top', 
          caption = "Count of Waze alerts by state and type, for the year from 2017-04-01 to 2018-03-30",
          options = list('pageLength' = 25)) %>%
          formatRound('count', digits = 0)
```

To do: same, but by month, counting distinct 5-min time intervals. Following chunks not complete. See Waze_Accident_Counts.Rmd for count of accidents by state.


  
  