---
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
      smooth_scroll: yes
author: "Safety Data Initiative | Volpe"
title: "Bellevue Waze Hot Spot Mapping"
---

<!-- saved from url=(0014)about:internet -->
<!-- This needs to be manually added to the second line of the html -->

```{r setup, echo = F, message = F, warning=F}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = F)
# Setup ----
# If you don't have these packages: install.packages(c("maps", "sp", "rgdal", "rgeos", "tidyverse")) 
# ggmap: want development version: devtools::install_github("dkahle/ggmap")
library(maps)
library(sp)
library(rgdal)
library(rgeos)
library(ggmap)
# library(spatstat) # for ppp density estimation. devtools::install_github('spatstat/spatstat')
library(tidyverse)
library(kableExtra)
library(DT)


codeloc <- "~/SDI_Waze"
source(file.path(codeloc, 'utility/get_packages.R'))
teambucket <- "s3://prod-sdc-sdi-911061262852-us-east-1-bucket"

user <- paste0( "/home/", system("whoami", intern = TRUE)) #the user directory to use
localdir <- paste0(user, "/workingdata") # full path for readOGR
state = "WA"
wazedir <- file.path(localdir, state) # has State_Year-mo.RData files. Grab from S3 if necessary

setwd(localdir) 

# read functions
source(file.path(codeloc, 'utility/wazefunctions.R'))
# Project to Albers equal area conic 102008. Check comparision wiht USGS version, WKID: 102039
proj <- showP4(showWKT("+init=epsg:102008"))
```

## Overview

Mapping of compiled Waze events from April 2017 to November 2018 in Bellevue, WA.
Events within 6 miles of the center of Bellevue were included in this initial work, and the 'city' field in the Waze data was used to remove events where were labeled as being outside of Bellevue. One of the next steps will be to clip to the city boundary shapefile. In addition, the Bellevue road network has not yet been incorporated in these maps. A base layer derived from Open Street Maps is used for these maps.

```{r load_data}
# load data ----

use.tz <- "US/Pacific"

if(length(grep("^WA_Bellevue_Prep.RData$", dir(file.path(localdir, "WA", "Waze"))))==0){ # do prep loop if needed

# Compile events ----

loc = c(-122.155, 47.59) # Bellevue, WA

wazemonthfiles <- dir(wazedir)[grep(state, dir(wazedir))]
wazemonthfiles <- wazemonthfiles[grep("RData$", wazemonthfiles)]
# omit _Raw_
omit <- grep('_Raw_', wazemonthfiles)
wazemonthfiles = wazemonthfiles[-omit]
wazemonthfiles = wazemonthfiles[nchar(wazemonthfiles)==16]

# Empty objects to store output, will rbind at the end of loop
waze.ll = waze.ll.proj = w.all = w.all.proj = vector()
  
for(i in 1:length(wazemonthfiles)){ # Start loop; i = 1
    cat(wazemonthfiles[i], "\n")
  
    load(file.path(wazedir, wazemonthfiles[i])) # load dataframe mb
  
  # Make both Waze and EDT spatial data frame, only picking out relevant columns. w and e are SPDF of Waze and EDT data, for this month
  w <- SpatialPointsDataFrame(mb[c("lon", "lat")], mb, 
                              proj4string = CRS("+proj=longlat +datum=WGS84"))  # from monthbind of Waze data, make sure it is a SPDF
  
  w <- spTransform(w, CRS(proj))
  # Subset to area of interest only ---- 
  zoomll = data.frame(Lon = loc[1], Lat = loc[2])
  
  zoomll = SpatialPoints(zoomll, 
                         proj4string = CRS("+proj=longlat +datum=WGS84"))
  
  zoomll <- spTransform(zoomll, CRS(proj))
  
  zoom_box = gBuffer(zoomll, width = 9654) # 6 mile circle around Bellevue
  
  zoom_box.ll = spTransform(zoom_box, CRS("+proj=longlat +datum=WGS84")) # Back to lat long for mapping
  
  # Step through Waze data to clip to special event buffer and day
  ws = gIntersects(zoom_box, w, byid = T)
  ws = w[ws[,1] == TRUE,] # Make ws as a spatial points data frame of Waze events, for points intersecting the buffer, in this month
  ws.ll = spTransform(ws, CRS("+proj=longlat +datum=WGS84")) # Back to lat long for mapping
 
  # Get data frames for plotting. wp and ep are data.frames of ws and es, with projected coordinates
  wp <- data.frame(ws.ll@data, lat = coordinates(ws)[,2], lon = coordinates(ws)[,1])
  
  ll <- data.frame(lat = coordinates(ws.ll)[,2], lon = coordinates(ws.ll)[,1], alert_type = ws.ll$alert_type, time = ws.ll$time, roadclass = ws.ll$road_type)
  
  ll.proj <- data.frame(lat = coordinates(ws)[,2], lon = coordinates(ws)[,1], alert_type = ws$alert_type, time = ws$time, roadclass = ws$road_type)
  
  # Compile
  w.all <- rbind(w.all, data.frame(ws.ll@coords, ws.ll@data))
  w.all.proj <- rbind(w.all.proj, data.frame(ws@coords, ws@data))
  
  waze.ll = rbind(waze.ll, ll)
  waze.ll.proj = rbind(waze.ll.proj, ll.proj)
 
} # End compile Waze event loop


# Save for plotting (will need to expand filename if doing multiple special events in a month) 
# w.all     Waze events for selected area, as spatial points data frame, with all variables
# waze.ll   Waze events for selected area, as spatial points data frame, with only alert type, road class, and lat long
# zoom_box  Bounding area for the geographic zoom
# zoomll    Lat/long of special event location at center of bounding area

# *.ll versions indicate represented as lat long instead of projected in Albers Equal Area; this makes mapping with ggmap more convenient for now. *.proj is projected versions

fn = paste(state, "Bellevue_Prep.RData", sep="_")

save(list = c("zoom_box", "zoomll",
              "w.all", "w.all.proj",
              "waze.ll", "waze.ll.proj",
              "zoom_box.ll"),
     file = file.path(localdir, "WA", "Waze",
                      fn))

fn.export = paste(state, 
                "Bellevue_Prep_Export.RData", sep="_")

save(list = c("zoom_box", "zoomll",
              "waze.ll", "waze.ll.proj",
              "zoom_box.ll"),
     file = file.path(localdir, "WA", "Waze",
                      fn.export))


# Save to S3
zipname = paste0('Hot_Spot_Mapping_Bellevue_prep_', Sys.Date(), '.zip')

system(paste('zip -j', file.path('~/workingdata', zipname),
             file.path(localdir, "WA", fn.export))
)

system(paste(
  'aws s3 cp',
  file.path('~/workingdata', zipname),
  file.path(teambucket, 'export_requests', zipname)
))
} else { # read already prepped data
  load(file.path(localdir, "WA", "Waze", "WA_Bellevue_Prep.RData"))
}

w.all <- w.all %>% filter(city == "Bellevue" | city == "Bellevue, WA" | is.na(city))

```

```{r plot_prep}
# Get a map for plotting.
if(length(grep("WA_Bellevue_Basemaps.RData", dir(file.path(localdir, "WA")))) == 0){
  map_terrain_14 <- get_stamenmap(bbox = as.vector(bbox(zoom_box.ll)), maptype = 'terrain', zoom = 14)
  map_toner_hybrid_14 <- get_stamenmap(bbox = as.vector(bbox(zoom_box.ll)), maptype = 'toner-hybrid', zoom = 14)
  map_toner_14 <- get_stamenmap(bbox = as.vector(bbox(zoom_box.ll)), maptype = 'toner', zoom = 14)
  save(list = c("map_terrain_14", "map_toner_hybrid_14", "map_toner_14"),
       file = file.path(localdir, "WA", "WA_Bellevue_Basemaps.RData"))
  } else { 
  load(file.path(localdir, "WA", "WA_Bellevue_Basemaps.RData"))
}

```

First, looking at frequency of all Waze incident reports by street name. Of all `r format(nrow(w.all), big.mark = ",")` Waze incident reports, `r format(nrow(w.all %>% filter(is.na(street))), big.mark = ",")` (`r round(100*nrow(w.all %>% filter(is.na(street)))/nrow(w.all), 2)`%) had no value provided for 'street' in the data. Of the `r format(nrow(w.all %>% filter(alert_type == "ACCIDENT")), big.mark = ",")` crash reports,  `r format(nrow(w.all %>% filter(is.na(street) & alert_type == "ACCIDENT")), big.mark = ",")` (`r round(100*nrow(w.all %>% filter(is.na(street) & alert_type == "ACCIDENT"))/nrow(w.all),2)`%) had no value provided for 'street'.


```{r street_table_hwy}
waze.ll$Date = format(waze.ll$time, "%Y-%m-%d")
waze.ll$YM = format(waze.ll$time, "%Y-%m")

# table(waze.ll %>% filter(alert_type =="ACCIDENT") %>% select(YM))
# ggplot(waze.ll %>% filter(alert_type == "ACCIDENT" & YM != "2017-03")) +
#   geom_histogram(aes(x = YM), stat="count") +
#   ylab("Count of Waze crash reports") + xlab("Year-month") +
#   ggtitle("Count of Waze crash reports by month within 6-mile buffer \n
#           of Bellevue, WA city center")

nametable <- as.data.frame(sort(table(as.factor(w.all$street)), decreasing = T))
colnames(nametable) = c("Street", "Frequency")
kable(nametable[1:10,], 
      caption = "Ten most frequent street names for all Waze incident reports in study period",
      format.args = list(big.mark = ",")) %>% 
  kable_styling(bootstrap_options = c('striped','hover'))
```

In the Waze data, there are two ways to make an initial filter to omit roads which are not in the jurisdiction of Bellevue: to filter by street names, and also to filter by road class. Waze uses their own road classification system; interstates are all class '3' in the Waze system.

Here we filter the Waze data by excluding I-405, I-90, as well as SR-520 (assuming this is not in Bellevue jurisdiction), and by omitting any additional events on class 3 roads. Filtering by name effectively removes nearly all class 3 roads. Some Waze events missing a street name and with an incorrect road class may still appear on the interstates after this filtering process; a GIS approach to snap Waze events to road segments will also be necessary.

```{r street_table_nohwy}
to_omit = c("I-405", "I-90", "SR-520", "I-5")

omits = vector()
for(i in to_omit){
  omits = c(omits, grep(paste0("^", i), w.all$street))
}
w.nohwy = w.all[is.na(match(1:nrow(w.all), omits)),]

# also omit by road_type = 3
w.nohwy = w.nohwy %>% filter(road_type != "3")

nametable <- as.data.frame(sort(table(as.factor(w.nohwy$street)), decreasing = T))
colnames(nametable) = c("Street", "Frequency")
kable(nametable[1:10,], 
      caption = "Ten most frequent street names for all Waze incident reports, omitting highways, in study period",
      format.args = list(big.mark = ",")) %>% 
  kable_styling(bootstrap_options = c('striped','hover'))
```

The Waze incidents remaining after filtering out highway events are summarized in the table below by alert type and subtype. Not all alerts have a subtype provided.

```{r summary_table_nohwy}
sum_count = w.nohwy %>%
  group_by(alert_type, sub_type) %>%
  summarize(count = n(),
            pct = round(100*count / nrow(w.nohwy), 2))

colnames(sum_count) = c("Alert type", "Subtype", "Count", "Percent of all alerts")
datatable(sum_count, caption = "Count of Waze incident repots by type and subtype, April 2017 to November 2018",
          rownames = F) %>% 
  formatCurrency(3, currency = "", digits = 0)
  
```


## Heat maps {.tabset .tabset-pills}

The following heat maps show 2-D kernel density estimates of incident frequency. Warm colors indicate high frequency, cool colors show lower frequency. Maps are shown with and without individual incidents plotted for clarity. 

### All Waze crash reports

`r format(nrow(w.all %>% filter(alert_type=="ACCIDENT")), big.mark=",")` Waze crash reports.

```{r Plotting_1, fig.width = 8}
  lx = w.all %>% select(lon, lat, alert_type, sub_type, road_type, time) %>% filter(alert_type == "ACCIDENT")

  lx.nohwy = w.nohwy %>% select(lon, lat, alert_type, sub_type, road_type, time) %>% filter(alert_type == "ACCIDENT")

  map1 <- ggmap(map_toner_hybrid_14, extent = 'device') + 
    stat_density2d(data = lx, aes(x = lon, y = lat,
                                  fill = ..level.., alpha = ..level..),
                   size = 0.01, bins = 8, geom = 'polygon',
                   h = 0.01) # manually set bandwith with h
  
  map1 + ggtitle(paste0(" Waze event density"))  +
          scale_fill_gradient(low = "blue", high = "red", 
                              guide_legend(title = "Event density")) +
          scale_alpha(range = c(0.1, 0.8), guide = FALSE) 
  
  # 2. With points
  map2 = map1 + 
    geom_point(data = lx, 
               aes(x = lon, y = lat), color = scales::alpha("midnightblue", 0.2), 
               pch = "+", cex = 3, stroke = 2) 
  
  map2 + ggtitle(paste0(" Waze crash report density"))  +
          scale_fill_gradient(low = "blue", high = "red", 
                              guide_legend(title = "Event density")) +
          scale_alpha(range = c(0.1, 0.8), guide = FALSE) 

```

### Waze crash reports, omitting highways

`r format(nrow(w.nohwy %>% filter(alert_type=="ACCIDENT")), big.mark=",")` Waze crash reports, after filtering out highways.

```{r Plotting_2, fig.width=8}

  map.nohwy <- ggmap(map_toner_hybrid_14, extent = 'device') + 
    stat_density2d(data = lx.nohwy, aes(x = lon, y = lat,
                                  fill = ..level.., alpha = ..level..),
                   size = 0.01, bins = 8, geom = 'polygon',
                   h = 0.01)
  
  map.nohwy + ggtitle(" Waze crash report density (omitting highways)")  +
          scale_fill_gradient(low = "blue", high = "red", 
                              guide_legend(title = "Crash density")) +
          scale_alpha(range = c(0.1, 0.8), guide = FALSE)
  

  # 2. With points
  map.nohwy.points = map.nohwy + 
    geom_point(data = lx.nohwy, 
               aes(x = lon, y = lat), color = scales::alpha("midnightblue", 0.2), 
               pch = "+", cex = 3, stroke = 2) 
  
  map.nohwy.points + ggtitle(paste0(" Waze crash report density (omitting highways)"))  +
          scale_fill_gradient(low = "blue", high = "red", 
                              guide_legend(title = "Event density")) +
          scale_alpha(range = c(0.1, 0.8), guide = FALSE) 
```  
  
