---
title: "Package installation on SDC"
output: html_document
---

Issue originally with `install.packages('tidyverse')` failing. Traces back to an issue with `openssl`, a necessary package for a number of purposes, including for `aws.s3`.

SDC instance is Linux, running Redhat OS. Openssl is installed, and can check from the terminal:

```
[daniel@ip-172-18-33-183 ~]$ openssl version -a
OpenSSL 1.0.2k-fips  26 Jan 2017
built on: reproducible build, date unspecified
platform: linux-x86_64
options:  bn(64,64) md2(int) rc4(16x,int) des(idx,cisc,16,int) idea(int) blowfish(idx)
compiler: gcc -I. -I.. -I../include  -fPIC -DOPENSSL_PIC -DZLIB -DOPENSSL_THREADS -D_REENTRANT -DDSO_DLFCN -DHAVE_DLFCN_H -DKRB5_MIT -m64 -DL_ENDIAN -Wall -O2 -g -pipe -Wall -Wp,-D_FORTIFY_SOURCE=2 -fexceptions -fstack-protector --param=ssp-buffer-size=4 -m64 -mtune=generic -Wa,--noexecstack -DPURIFY -DOPENSSL_IA32_SSE2 -DOPENSSL_BN_ASM_MONT -DOPENSSL_BN_ASM_MONT5 -DOPENSSL_BN_ASM_GF2m -DRC4_ASM -DSHA1_ASM -DSHA256_ASM -DSHA512_ASM -DMD5_ASM -DAES_ASM -DVPAES_ASM -DBSAES_ASM -DWHIRLPOOL_ASM -DGHASH_ASM -DECP_NISTZ256_ASM
OPENSSLDIR: "/etc/pki/tls"
engines:  rdrand dynamic
```
Can also confirm with `yum info openssl`, looks good.

Ok, so we have openssl accessible in the operating system. From Installation notes for openssl:

"On Fedora, CentOS or RHEL use openssl-devel:"
```
sudo yum install openssl-devel
```

We can't do this as a user, since not in the sudo file.. can in root, in Jupyter notebook terminal:

```
ec2-user@ip-172-18-33-183 home]$ sudo yum install openssl-devel
Loaded plugins: priorities, update-motd, upgrade-helper
amzn-main                                                                                                          | 2.1 kB  00:00:00
amzn-updates                                                                                                       | 2.5 kB  00:00:00
Package 1:openssl-devel-1.0.2k-8.106.amzn1.x86_64 already installed and latest version
Nothing to do
```

But this doesn't solve the issue! 

Try installing the R package from GitHub directly! **This works**. 
Same problem with `selectr`. 

```{r eval=FALSE, echo=T}

if(length(grep("devtools", (.packages(all.available=T))))==0) install.packages("devtools", dependencies =TRUE)

devtools::install_github("jeroen/openssl")

devtools::install_github("sjp/selectr")

```

Now load the packages we want:

```{r, eval=F}

loadpacks <- c("httr", 
               "rvest", # was still failing, needed selectr also
               "aws.s3",
               "tidyverse" # do this after rvest
               )
for(i in loadpacks){if(length(grep(i, (.packages(all.available=T))))==0) install.packages(i, dependencies =TRUE)}

```

# GIS Packages on SDC
## `rgeos` and `rgdal`

Both `rgeos` and `rgdal` depend some tools that need to be installed as root:

References [1](https://askubuntu.com/questions/312294/once-installed-geos-library-c-and-c-and-then-trying-to-install-rgeos-packa) and [2](https://stackoverflow.com/questions/21683138/unable-to-install-rgdal-and-rgeos-r-libraries-on-red-hat-linux#24866775)

```
sudo yum install geos geos-devel
```
Now try again:

```{r, eval = F}
install.packages(c("rgeos", "rgdal"), dep=T, type = "source")
```

This fails with the error:
`libgdal.so.20: cannot open shared object file: No such file or directory`

This exists, and is in `/usr/local/lib`, see
```
whereis 'libgdal.so.20'
```

Try again, by loading packages from the Extra Packages for Enterprise Linux repository (EPEL):

https://aws.amazon.com/premiumsupport/knowledge-center/ec2-enable-epel/

Now try to install `gdal`:

```
sudo yum --enablerepo=epel install -y gdal proj proj-devel gdal-devel proj-epsg proj-nad

```

See that gdal is installed with:

```
gdalinfo --version
```

Still doesn't work. Try updating the dynamic library symbolic links
https://stackoverflow.com/questions/40305095/cannot-install-rgdal-package-in-r-on-rhel6-unable-to-load-shared-object-rgdal-s


```
echo "/usr/local/lib" >> /etc/ld.so.conf.d/R-dependencies-x86_64.conf
ldconfig
```
Do not have permission to write to /etc/ld.so.conf.d/, so run this instead:

```
sudo ldconfig -c "echo '/usr/local/lib' >> /etc/ld.so.conf.d/R-dependencies-x86_64.conf"

```


Try updating the PATH:

https://stackoverflow.com/questions/48823509/error-loading-rgdal

```
export LD_LIBRARY_PATH="/usr/bin/gdal/lib:$LD_LIBRARY_PATH"
ldd /home/daniel/R/x86_64-pc-linux-gnu-library/3.4/rgdal/libs/rgdal.so
```

Where is gdal-libs? This seems to be the problem, mentioned in several solutions on SO, but not available via yum. Look at what gdal packages are available:

```
yum repo-pkgs epel list gdal*
sudo yum search all "gdal"
```

No luck, try this:


```{r, eval=F}
install.packages("rgdal", method="curl", configure.args = c("--with-proj-include=/usr/bin/proj/include","--with-proj-lib=/usr/bin/proj/lib"))
```
Same problem as above.

Ok, trying to get gdal-libs installed by manually downloading from RedHat Package Manager:
```
wget http://mirror.nodesdirect.com/epel/7/x86_64/Packages/g/gdal-libs-1.11.4-3.el7.x86_64.rpm
sudo yum install gdal-libs-1.11.4-3.el7.x86_64.rpm

wget https://rpmfind.net/linux/epel/6/x86_64/Packages/c/cfitsio-3.240-3.el6.x86_64.rpm
```






Edit: nope, also doesnot work:

```
sudo yum -y update
sudo yum-config-manager --enable epel
sudo yum -y install make automake gcc gcc-c++ libcurl-devel proj-devel geos-devel
cd /tmp
curl -L http://download.osgeo.org/gdal/2.2.3/gdal-2.2.3.tar.gz | tar zxf -
cd gdal-2.3.3/
./configure --prefix=/usr/local --without-python
sudo make -j4
sudo make install
cd /usr/local
tar zcvf ~/gdal-2.0.0-amz1.tar.gz *
```
