---
title: "Special Events Visualization - Part 1"
author: "Ci (Jessie) Yang"
date: "`r paste('Last Updated:',format(Sys.time(), '%B %d, %Y'))`"
output:
  html_document: default
pdf_document:
  fig_caption: yes
fig_height: 5
fig_width: 8
---

This is an example of visuals that we can create visualizations for a specific location and buffer zone around the location where we have model outputs on.

```{r setup, include=FALSE}
#set options
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE)
options(scipen = 20)

codeloc <- "~/Github/SDI_Waze"
source(file.path(codeloc, 'utility/get_packages.R'))

library(tidyverse)
library(sp)
library(maps) # for mapping base layers
library(rgdal) # for readOGR(), needed for reading in ArcM shapefiles
library(foreach)
library(doParallel)
library(rgeos) #gintersection
library(lubridate)
library(ggplot2)
library(extrafont)
library(splus2R)

localdir <- "C:/Users/Jessie.Yang.CTR/Downloads/OST/Waze project/Special Events"
# localdir <- "/home/daniel/workingdata/" # full path for readOGR, Jessie don't have this folder.
# edtdir <- normalizePath(file.path(localdir, "EDT"))
# wazedir <- "~/tempout" # has State_Year-mo.RData files. Grab from S3 if necessary
wazedir <- "//vntscex.local/DFS/Projects/PROJ-OS62A1/SDI Waze Phase 2"
output.loc <- file.path(wazedir, "WazeEDT Pilot Phase1 Archive Incomplete/SDI/Model_Output")
data.loc <- file.path(wazedir, "WazeEDT Pilot Phase1 Archive Incomplete/SDI/Data")

teambucket <- "s3://prod-sdc-sdi-911061262852-us-east-1-bucket"

source(file.path(codeloc, "utility/wazefunctions.R")) 

states = c("CT", "UT", "VA", "MD")

# Time zone picker:
tzs <- data.frame(states, 
                  tz = c("US/Eastern",
                         "US/Mountain",
                         "US/Eastern",
                         "US/Eastern"),
                  stringsAsFactors = F)


# Project to Albers equal area conic 102008. Check comparision with USGS version, WKID: 102039
proj <- showP4(showWKT("+init=epsg:102008"))
# USGS version, used for producing hexagons: 
proj.USGS <- "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=23 +lon_0=-96 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0"

load(file = paste0(wazedir,"/Data/SpecialEvents/SpecialEvents_MD_AprilToSept_2017.Rdata"))

# ggplot format
theme_set(theme_bw(base_size = 12) +
            theme(plot.background = element_blank(),
                  panel.grid.major = element_blank(),
                  panel.grid.minor = element_blank(),
                  panel.border = element_blank(),
                  text = element_text(family="Gill Sans MT", size=12,colour = "#666666"),
                  legend.text = element_text(family="Gill Sans MT", size=10, colour = "#666666"),
                  plot.title = element_text(size = rel(1),
                                            colour = "#666666", vjust=2, hjust=0.5, face="bold",family="Gill Sans MT"),
                  strip.text.x = element_text(size = 11,colour = "#666666",family="Gill Sans MT"),
                  axis.text.x=element_text(size=11, angle = 0, hjust = 0.5,family="Gill Sans MT",colour = "#666666"),
                  axis.text.y=element_text(size=11, angle = 0, hjust = 1,family="Gill Sans MT",colour = "#666666"))
)

# function to convert military time to HH:MM
mil_to_hm <- function(x){
  format(strptime(substr(as.POSIXct(sprintf("%04.0f", x), 
                                    format="%H%M"), 
                         12, 16), '%H:%M')
         , '%H:%M')
}

# function to convert HH:MM to decimals
hm_to_num <- function(x){
  sapply(strsplit(x,":"),
  function(x) {
    x <- as.numeric(x)
    x[1]+x[2]/60
    }
)
}
```

# Time Series Plots

We used the AllModel30.csv output for this analysis.

Metrics we use include:

* Number of EDT Accidents (Model Dependent Variable)
* Number of Waze Accidents
* Number of Waze Jam
* Number of Hazard On Shoulder (nHazardOnShoulder)
* Number of Hazard On Road (nHazardOnRoad)
* Number of Hazard Weather (nHazardWeather) - is it # of hazard or weather? According to Michelle, this should be one of the three categories of Waze event (Accidents, Jam, and Hazard/Weather). However, in the dataset, nHazardWeather is smaller than sum of the sub-categories (e.g., nHazardOnShoulder, nHazardOnRoad). This needs further investigation.
* Probability of NonCrash and Crash - real prob from RF model (needs further investigation/decision making)
The probability has converted to based on the threshold ~ 0.2 using the built-in function in RF.

Following visualizations are for location "SE1" and 1 mile buffer zone.

To do 8/23/2018:

* total/average/max (point/line) number of accidents/jams for every day (done)
    * plot on top of each other for the NoEvents with a different color. Use other colors for a type of special events. 4/23 (4/16 easter), 5/21 (5/29 memorial day, 5/14 mother's day), 7/02 weekends near holidays. 
    * Do both 3-miles and 1-mile. Removing zeros. 
    * Panels by month.
    * add shadebox for event duration
* add function for visuals
* number of accidents per gridcell per hour
* boxplot of accidents over 42 grid cells. (done)
* look at hour aggregation
* heatmap for each hour
* smaller buffer zone, 0.25, 0.5, 1 mile radius 
    + update SpecialEventsExpand dataset (done)
    + update the analysis and visuals
* Easten shore beach - Bay Bridge area (7/4, 9/4 Labor Day, 3lanes both ways) for a typical holiday pattern
* follow up with Erika tomorrow

```{r data Process, include = FALSE}
# To recover the grid_ids for 1 mile buffer of location SE1
loc = "SE1"
# buf = 1

# Get all combinations of grid_id, hour, and day
alldays = seq(from = min(AllModel30_sub$date), to = max(AllModel30_sub$date), by = "days") # get all days during April - Sep
days = alldays[as.numeric(format(alldays, format = "%u")) %in% unique(SpecialEventsExpand$DayofWeekN)] # get all Sundays, Saturdays, Tuesdays, and Wednesdays that had special events
daterange = days

# Numeric variables if missing, can be replaced using zeros
col.names <- c("Obs","nWazeAccident","nWazeJam","nHazardOnShoulder","nHazardOnRoad","nHazardWeather")

# Function to create GridData joined with Special events, the special events are linked to date, not hours.
GridDataSE <- function(loc,buf,daterange,col.names){
  # recover the grid_ids from column GRID_ID
  grid_id <- unlist(strsplit(unique(SpecialEventsExpand$GRID_ID[SpecialEventsExpand$Buffer_Miles == buf & SpecialEventsExpand$Location.ID == loc]), split = ",")
  ) # an example of extracting grib_id of the largest buffer
  
  # all possible combinations
  gridcom <- expand.grid(Location.ID = loc, GRID_ID = grid_id, hour = c(0:23), date = daterange) # for example, if daterange include 105 days, then row number 105840 = 105 days*24 hour*42 grid_id
  
  # create new variables
  gridcom <- gridcom %>% mutate(day = yday(date), # convert day to date
                                weekday = as.numeric(format(date, format = "%u")), # convert weekday
                                DayofWeek = weekdays(date) #wday(date, label = T) returns a full week of day name
  )
  
  # grid cells in the model
  grid_cells_model <- unique(AllModel30_sub$GRID_ID)
  
  # left_join the Event Type to the model output data
  dt <- gridcom %>% left_join(AllModel30_sub, by = c("GRID_ID", "day", "hour","date", "weekday")) %>% mutate_if(colnames(.) %in% col.names,funs(replace(., which(is.na(.)), 0))) %>% left_join(SpecialEvents[,c("Day","EventType","StartTime","EndTime")], by =  c("day" = "Day")) %>% mutate_if(colnames(.) %in% c("EventType"),funs(replace(., which(is.na(.)), "NoEvent"))) %>% mutate(Grid_In_Model = ifelse(GRID_ID %in% grid_cells_model, "Yes", "No"))
  dt
}

# # Function to create time series by day of week, month, and date
# ts_day_month <- function(buf, stat ){
#   
# }

```

## Hourly By Event Type

### 3 mile buffer zone - Averages
```{r eval = F}
buf = 3
dt <- GridDataSE(loc,buf,daterange,col.names)

plotAvgJamMonth <- function(dt, weekday){
            # Mean
          dt_Date <- dt %>% group_by(Location.ID, date, hour) %>% summarize(nWazeJam = mean(nWazeJam),
                                                                         Obs = mean(Obs),
                                                                         nWazeAccident = mean(nWazeAccident),
                                                                         nHazardOnShoulder = mean(nHazardOnShoulder),
                                                                         nHazardOnRoad = mean(nHazardOnRoad),
                                                                         nHazardWeather = mean(nHazardWeather)
          ) %>% mutate(DayofWeek = weekdays(date), Date = paste(date, DayofWeek), Month = month(date)) %>%  left_join(unique(dt[, c("date","EventType","StartTime","EndTime")]), by = "date") %>% mutate(background_col = ifelse(EventType == "NoEvent", NA,ifelse(EventType == "Football","#E69F00", "#56B4E9")))
          
          dt_Date$Month <- factor(dt_Date$Month) 
          levels(dt_Date$Month) <- list("April" = 4, "May" = 5, "June" = 6, "July" = 7, "August" = 8, "September" = 9)
          
          # Military time to regular HH:MM, and then to decimals.
          dt_Date$StartTime <- hm_to_num(mil_to_hm(dt_Date$StartTime))
          dt_Date$EndTime <- hm_to_num(mil_to_hm(dt_Date$EndTime))
          
          dt_visual <-  dt_Date %>% filter(DayofWeek == weekday)
          
          ggplot(dt_visual, aes(x = hour, y = nWazeJam, color = EventType, group = date)) + 
            # geom_point(alpha = 0.5, color = "red") +
            geom_line(alpha = 0.5) +
            geom_rect(data = dt_visual,aes(xmin = ave(dt_visual$StartTime, dt_visual$date, FUN = function(x) mean(x)),xmax = ave(dt_visual$EndTime, dt_visual$date, FUN = function(x) mean(x)), ymin = -Inf,ymax=Inf), fill = dt_visual$background_col, alpha= 0.2/100, inherit.aes = FALSE) +
            facet_wrap(~ Month) +
            ylab("Average Waze Jam") +
            xlab("Hour of Day") +
            scale_color_manual(name = NULL, values=c("NoEvent" = "#999999", "Football" = "#E69F00", "Concert" = "#56B4E9")) +
            ggtitle(paste("Fedex Field,", buf,"Mile Buffer,", weekday)) 
}

weekday = "Sunday"
plotAvgJamMonth(dt, weekday)


# weekday = "Saturday"
# plotAvgJamMonth(dt, weekday)
# 
# weekday = "Tuesday"
# plotAvgJamMonth(dt, weekday)
# 
# weekday = "Wednesday"
# plotAvgJamMonth(dt, weekday)
```

### 3 mile buffer zone - Averages without zeros
```{r, eval = F}
buf = 3
dt <- GridDataSE(loc,buf,daterange,col.names)

# Mean
dt_Date <- dt %>% group_by(Location.ID, date, hour) %>% summarize(nWazeJam = mean(nWazeJam),
                                                               Obs = mean(Obs),
                                                               nWazeAccident = mean(nWazeAccident),
                                                               nHazardOnShoulder = mean(nHazardOnShoulder),
                                                               nHazardOnRoad = mean(nHazardOnRoad),
                                                               nHazardWeather = mean(nHazardWeather)
) %>% mutate(DayofWeek = weekdays(date), Date = paste(date, DayofWeek), Month = month(date)) %>%  left_join(unique(dt[, c("date","EventType","StartTime","EndTime")]), by = "date") %>% mutate(background_col = ifelse(EventType == "NoEvent", NA,ifelse(EventType == "Football","#E69F00", "#56B4E9"))) %>% mutate_at(.vars = vars(Obs, nWazeAccident, nWazeJam,nHazardOnShoulder,nHazardOnRoad),.funs = funs(replace(., which(. == 0), NA)))

dt_Date$Month <- factor(dt_Date$Month) 
levels(dt_Date$Month) <- list("April" = 4, "May" = 5, "June" = 6, "July" = 7, "August" = 8, "September" = 9)

# Military time to regular HH:MM, and then to decimals.
dt_Date$StartTime <- hm_to_num(mil_to_hm(dt_Date$StartTime))
dt_Date$EndTime <- hm_to_num(mil_to_hm(dt_Date$EndTime))

weekday = "Sunday"
dt_visual <-  dt_Date %>% filter(DayofWeek == weekday)

ggplot(dt_visual, aes(x = hour, y = nWazeJam, color = EventType, group = date)) + 
  # geom_point(alpha = 0.5, color = "black") +
  geom_line(alpha = 0.5) +
  geom_rect(data = dt_visual,aes(xmin = ave(dt_visual$StartTime, dt_visual$date, FUN = function(x) mean(x)),xmax = ave(dt_visual$EndTime, dt_visual$date, FUN = function(x) mean(x)), ymin = -Inf,ymax=Inf), fill = dt_visual$background_col, alpha= 0.2/100, inherit.aes = FALSE) +
  facet_wrap(~ Month) +
  ylab("Average Waze Jam") +
  xlab("Hour of Day") +
  scale_color_manual(name = NULL, values=c("NoEvent" = "#999999", "Football" = "#E69F00", "Concert" = "#56B4E9")) +
  ggtitle(paste("Fedex Field,", buf,"Mile Buffer,", weekday)) 
```

### 1 mile buffer zone - Averages
```{r eval = F}
buf = 1
dt <- GridDataSE(loc,buf,daterange,col.names)

weekday = "Sunday"
plotAvgJamMonth(dt, weekday)

# weekday = "Saturday"
# plotAvgJamMonth(dt, weekday)
# 
# weekday = "Tuesday"
# plotAvgJamMonth(dt, weekday)
# 
# weekday = "Wednesday"
# plotAvgJamMonth(dt, weekday)
```
### 0.5 mile buffer zone - Averages
```{r eval = F}
buf = 0.5
dt <- GridDataSE(loc,buf,daterange,col.names)

weekday = "Sunday"
plotAvgJamMonth(dt, weekday)

# weekday = "Saturday"
# plotAvgJamMonth(dt, weekday)
# 
# weekday = "Tuesday"
# plotAvgJamMonth(dt, weekday)
# 
# weekday = "Wednesday"
# plotAvgJamMonth(dt, weekday)
```

### 3 mile buffer zone - Sum Value
```{r}
buf = 3
dt <- GridDataSE(loc,buf,daterange,col.names)

plotSumJamMonth <- function(dt, weekday) {
        # Total
      dt_Date <- dt %>% group_by(Location.ID, date, hour) %>% summarize(nWazeJam = sum(nWazeJam),
                                                                     Obs = sum(Obs),
                                                                     nWazeAccident = sum(nWazeAccident),
                                                                     nHazardOnShoulder = sum(nHazardOnShoulder),
                                                                     nHazardOnRoad = sum(nHazardOnRoad),
                                                                     nHazardWeather = sum(nHazardWeather)
      ) %>% mutate(DayofWeek = weekdays(date), Date = paste(date, DayofWeek), Month = month(date)) %>%  left_join(unique(dt[, c("date","EventType","StartTime","EndTime")]), by = "date") %>% mutate(background_col = ifelse(EventType == "NoEvent", NA,ifelse(EventType == "Football","#E69F00", "#56B4E9")))
      
      dt_Date$Month <- factor(dt_Date$Month) 
      levels(dt_Date$Month) <- list("April" = 4, "May" = 5, "June" = 6, "July" = 7, "August" = 8, "September" = 9)
      
      # Military time to regular HH:MM, and then to decimals.
      dt_Date$StartTime <- hm_to_num(mil_to_hm(dt_Date$StartTime))
      dt_Date$EndTime <- hm_to_num(mil_to_hm(dt_Date$EndTime))
      
      weekday = "Sunday"
      dt_visual <-  dt_Date %>% filter(DayofWeek == weekday)
      
      ggplot(dt_visual, aes(x = hour, y = nWazeJam, color = EventType, group = date)) + 
        # geom_point(alpha = 0.5, color = "red") +
        geom_line(alpha = 0.5) +
        geom_rect(data = dt_visual,aes(xmin = ave(dt_visual$StartTime, dt_visual$date, FUN = function(x) mean(x)),xmax = ave(dt_visual$EndTime, dt_visual$date, FUN = function(x) mean(x)), ymin = -Inf,ymax=Inf), fill = dt_visual$background_col, alpha= 0.2/100, inherit.aes = FALSE) +
        facet_wrap(~ Month) +
        ylab("Total Waze Jam") +
        xlab("Hour of Day") +
        scale_color_manual(name = NULL, values=c("NoEvent" = "#999999", "Football" = "#E69F00", "Concert" = "#56B4E9")) +
        ggtitle(paste("Fedex Field,", buf,"Mile Buffer,", weekday)) 
}
plotSumCrashMonth <- function(dt, weekday) {
        # Total
      dt_Date <- dt %>% group_by(Location.ID, date, hour) %>% summarize(nWazeJam = sum(nWazeJam),
                                                                     Obs = sum(Obs),
                                                                     nWazeAccident = sum(nWazeAccident),
                                                                     nHazardOnShoulder = sum(nHazardOnShoulder),
                                                                     nHazardOnRoad = sum(nHazardOnRoad),
                                                                     nHazardWeather = sum(nHazardWeather)
      ) %>% mutate(DayofWeek = weekdays(date), Date = paste(date, DayofWeek), Month = month(date)) %>%  left_join(unique(dt[, c("date","EventType","StartTime","EndTime")]), by = "date") %>% mutate(background_col = ifelse(EventType == "NoEvent", NA,ifelse(EventType == "Football","#E69F00", "#56B4E9")))
      
      dt_Date$Month <- factor(dt_Date$Month) 
      levels(dt_Date$Month) <- list("April" = 4, "May" = 5, "June" = 6, "July" = 7, "August" = 8, "September" = 9)
      
      # Military time to regular HH:MM, and then to decimals.
      dt_Date$StartTime <- hm_to_num(mil_to_hm(dt_Date$StartTime))
      dt_Date$EndTime <- hm_to_num(mil_to_hm(dt_Date$EndTime))
      
      weekday = "Sunday"
      dt_visual <-  dt_Date %>% filter(DayofWeek == weekday)
      
      ggplot(dt_visual, aes(x = hour, y = nWazeAccident, color = EventType, group = date)) + 
        # geom_point(alpha = 0.5, color = "red") +
        geom_line(alpha = 0.5) +
        geom_rect(data = dt_visual,aes(xmin = ave(dt_visual$StartTime, dt_visual$date, FUN = function(x) mean(x)),xmax = ave(dt_visual$EndTime, dt_visual$date, FUN = function(x) mean(x)), ymin = -Inf,ymax=Inf), fill = dt_visual$background_col, alpha= 0.2/100, inherit.aes = FALSE) +
        facet_wrap(~ Month) +
        ylab("Total Waze Accident") +
        xlab("Hour of Day") +
        scale_color_manual(name = NULL, values=c("NoEvent" = "#999999", "Football" = "#E69F00", "Concert" = "#56B4E9")) +
        ggtitle(paste("Fedex Field,", buf,"Mile Buffer,", weekday)) 
}
plotSumEDTMonth <- function(dt, weekday) {
        # Total
      dt_Date <- dt %>% group_by(Location.ID, date, hour) %>% summarize(nWazeJam = sum(nWazeJam),
                                                                     Obs = sum(Obs),
                                                                     nWazeAccident = sum(nWazeAccident),
                                                                     nHazardOnShoulder = sum(nHazardOnShoulder),
                                                                     nHazardOnRoad = sum(nHazardOnRoad),
                                                                     nHazardWeather = sum(nHazardWeather)
      ) %>% mutate(DayofWeek = weekdays(date), Date = paste(date, DayofWeek), Month = month(date)) %>%  left_join(unique(dt[, c("date","EventType","StartTime","EndTime")]), by = "date") %>% mutate(background_col = ifelse(EventType == "NoEvent", NA,ifelse(EventType == "Football","#E69F00", "#56B4E9")))
      
      dt_Date$Month <- factor(dt_Date$Month) 
      levels(dt_Date$Month) <- list("April" = 4, "May" = 5, "June" = 6, "July" = 7, "August" = 8, "September" = 9)
      
      # Military time to regular HH:MM, and then to decimals.
      dt_Date$StartTime <- hm_to_num(mil_to_hm(dt_Date$StartTime))
      dt_Date$EndTime <- hm_to_num(mil_to_hm(dt_Date$EndTime))
      
      weekday = "Sunday"
      dt_visual <-  dt_Date %>% filter(DayofWeek == weekday)
      
      ggplot(dt_visual, aes(x = hour, y = Obs, color = EventType, group = date)) + 
        # geom_point(alpha = 0.5, color = "red") +
        geom_line(alpha = 0.5) +
        geom_rect(data = dt_visual,aes(xmin = ave(dt_visual$StartTime, dt_visual$date, FUN = function(x) mean(x)),xmax = ave(dt_visual$EndTime, dt_visual$date, FUN = function(x) mean(x)), ymin = -Inf,ymax=Inf), fill = dt_visual$background_col, alpha= 0.2/100, inherit.aes = FALSE) +
        facet_wrap(~ Month) +
        ylab("Total EDT Accident") +
        xlab("Hour of Day") +
        scale_color_manual(name = NULL, values=c("NoEvent" = "#999999", "Football" = "#E69F00", "Concert" = "#56B4E9")) +
        ggtitle(paste("Fedex Field,", buf,"Mile Buffer,", weekday)) 
}
weekday = "Sunday"
plotSumJamMonth(dt, weekday)
plotSumCrashMonth(dt, weekday)
plotSumEDTMonth(dt, weekday)
```


### 1 mile buffer zone - Sum Value
```{r}
buf = 1
dt <- GridDataSE(loc,buf,daterange,col.names)

weekday = "Sunday"
plotSumJamMonth(dt, weekday)
plotSumCrashMonth(dt, weekday)
plotSumEDTMonth(dt, weekday)
```

### half mile buffer zone - Sum Value
```{r}
buf = 0.5
dt <- GridDataSE(loc,buf,daterange,col.names)

weekday = "Sunday"
plotSumJamMonth(dt, weekday)
plotSumCrashMonth(dt, weekday)
plotSumEDTMonth(dt, weekday)
```


### 3 mile buffer zone - Max Value
```{r eval = F}
buf = 3
dt <- GridDataSE(loc,buf,daterange,col.names)

plotMaxJamMonth <- function(dt, weekday){
  # Max
dt_Date <- dt %>% group_by(Location.ID, date, hour) %>% summarize(nWazeJam = max(nWazeJam),
                                                               Obs = max(Obs),
                                                               nWazeAccident = max(nWazeAccident),
                                                               nHazardOnShoulder = max(nHazardOnShoulder),
                                                               nHazardOnRoad = max(nHazardOnRoad),
                                                               nHazardWeather = max(nHazardWeather)
) %>% mutate(DayofWeek = weekdays(date), Date = paste(date, DayofWeek), Month = month(date)) %>%  left_join(unique(dt[, c("date","EventType","StartTime","EndTime")]), by = "date") %>% mutate(background_col = ifelse(EventType == "NoEvent", NA,ifelse(EventType == "Football","#E69F00", "#56B4E9")))

dt_Date$Month <- factor(dt_Date$Month) 
levels(dt_Date$Month) <- list("April" = 4, "May" = 5, "June" = 6, "July" = 7, "August" = 8, "September" = 9)

# Military time to regular HH:MM, and then to decimals.
dt_Date$StartTime <- hm_to_num(mil_to_hm(dt_Date$StartTime))
dt_Date$EndTime <- hm_to_num(mil_to_hm(dt_Date$EndTime))

weekday = "Sunday"
dt_visual <-  dt_Date %>% filter(DayofWeek == weekday)

ggplot(dt_visual, aes(x = hour, y = nWazeJam, color = EventType, group = date)) + 
  # geom_point(alpha = 0.5, color = "red") +
  geom_line(alpha = 0.5) +
  geom_rect(data = dt_visual,aes(xmin = ave(dt_visual$StartTime, dt_visual$date, FUN = function(x) unique(x)),xmax = ave(dt_visual$EndTime, dt_visual$date, FUN = function(x) mean(x)), ymin = -Inf,ymax=Inf), fill = dt_visual$background_col, alpha= 0.2/100, inherit.aes = FALSE) +
  facet_wrap(~ Month) +
  ylab("Max Waze Jam") +
  xlab("Hour of Day") +
  scale_color_manual(name = NULL, values=c("NoEvent" = "#999999", "Football" = "#E69F00", "Concert" = "#56B4E9")) +
  ggtitle(paste("Fedex Field,", buf,"Mile Buffer,", weekday)) 
}

weekday = "Sunday"
plotMaxJamMonth(dt, weekday)
```

### 1 mile buffer zone - Max Value
```{r eval = F}
buf = 1
dt <- GridDataSE(loc,buf,daterange,col.names)

weekday = "Sunday"
plotMaxJamMonth(dt, weekday)
```

### half mile buffer zone - Max Value
```{r eval = F}
buf = 0.5
dt <- GridDataSE(loc,buf,daterange,col.names)

weekday = "Sunday"
plotMaxJamMonth(dt, weekday)
```

### Hourly By Day of Week - 3 mile buffer

```{r, fig.width=12, fig.height=12}
buf = 3
dt <- GridDataSE(loc,buf,daterange,col.names)

plotSumJam <- function (dt, weekday) {
            # Mean
          dt_Date <- dt %>% group_by(Location.ID, date, hour) %>% summarize(nWazeJam = sum(nWazeJam),
                                                                         Obs = sum(Obs),
                                                                         nWazeAccident = sum(nWazeAccident),
                                                                         nHazardOnShoulder = sum(nHazardOnShoulder),
                                                                         nHazardOnRoad = sum(nHazardOnRoad),
                                                                         nHazardWeather = sum(nHazardWeather)
          ) %>% mutate(DayofWeek = weekdays(date), Date = paste(date, DayofWeek), Month = month(date)) %>%  left_join(unique(dt[, c("date","EventType","StartTime","EndTime")]), by = "date") %>% mutate(background_col = ifelse(EventType == "NoEvent", NA,ifelse(EventType == "Football","#E69F00",ifelse(EventType == "Concert", "#56B4E9", "chartreuse4" ))))
          
          dt_Date$Month <- factor(dt_Date$Month) 
          levels(dt_Date$Month) <- list("April" = 4, "May" = 5, "June" = 6, "July" = 7, "August" = 8, "September" = 9)
          
          # Military time to regular HH:MM, and then to decimals.
          dt_Date$StartTime <- hm_to_num(mil_to_hm(dt_Date$StartTime))
          dt_Date$EndTime <- hm_to_num(mil_to_hm(dt_Date$EndTime))
          
          dt_visual <-  dt_Date %>% filter(DayofWeek == weekday)
          
          ggplot(dt_visual, aes(x = hour, y = nWazeJam, color = EventType, group = date)) + 
            # geom_point(alpha = 0.5, color = "red") +
            geom_line(alpha = 0.5) +
            geom_rect(data = dt_visual,aes(xmin = ave(dt_visual$StartTime, dt_visual$date, FUN = function(x) mean(x)),xmax = ave(dt_visual$EndTime, dt_visual$date, FUN = function(x) mean(x)), ymin = -Inf,ymax=Inf), fill = dt_visual$background_col, alpha= 0.2/100, inherit.aes = FALSE) +
            facet_wrap(~ Date) +
            ylab("Total Waze Jam") +
            xlab("Hour of Day") +
            scale_color_manual(name = NULL, values=c("NoEvent" = "#999999", "Football" = "#E69F00", "Concert" = "#56B4E9", "Soccer" = "chartreuse4")) +
            ggtitle(paste("Fedex Field,", buf,"Mile Buffer,", weekday)) 
}

plotSumCrash <- function (dt, weekday) {
            # Mean
          dt_Date <- dt %>% group_by(Location.ID, date, hour) %>% summarize(nWazeJam = sum(nWazeJam),
                                                                         Obs = sum(Obs),
                                                                         nWazeAccident = sum(nWazeAccident),
                                                                         nHazardOnShoulder = sum(nHazardOnShoulder),
                                                                         nHazardOnRoad = sum(nHazardOnRoad),
                                                                         nHazardWeather = sum(nHazardWeather)
          ) %>% mutate(DayofWeek = weekdays(date), Date = paste(date, DayofWeek), Month = month(date)) %>%  left_join(unique(dt[, c("date","EventType","StartTime","EndTime")]), by = "date") %>% mutate(background_col = ifelse(EventType == "NoEvent", NA,ifelse(EventType == "Football","#E69F00",ifelse(EventType == "Concert", "#56B4E9", "chartreuse4" ))))
          
          dt_Date$Month <- factor(dt_Date$Month) 
          levels(dt_Date$Month) <- list("April" = 4, "May" = 5, "June" = 6, "July" = 7, "August" = 8, "September" = 9)
          
          # Military time to regular HH:MM, and then to decimals.
          dt_Date$StartTime <- hm_to_num(mil_to_hm(dt_Date$StartTime))
          dt_Date$EndTime <- hm_to_num(mil_to_hm(dt_Date$EndTime))
          
          dt_visual <-  dt_Date %>% filter(DayofWeek == weekday)
          
          ggplot(dt_visual, aes(x = hour, y = nWazeAccident, color = EventType, group = date)) + 
            # geom_point(alpha = 0.5, color = "red") +
            geom_line(alpha = 0.5) +
            geom_rect(data = dt_visual,aes(xmin = ave(dt_visual$StartTime, dt_visual$date, FUN = function(x) mean(x)),xmax = ave(dt_visual$EndTime, dt_visual$date, FUN = function(x) mean(x)), ymin = -Inf,ymax=Inf), fill = dt_visual$background_col, alpha= 0.2/100, inherit.aes = FALSE) +
            facet_wrap(~ Date) +
            ylab("Total Waze Accident") +
            xlab("Hour of Day") +
            scale_color_manual(name = NULL, values=c("NoEvent" = "#999999", "Football" = "#E69F00", "Concert" = "#56B4E9", "Soccer" = "chartreuse4")) +
            ggtitle(paste("Fedex Field,", buf,"Mile Buffer,", weekday)) 
}

plotSumEDT <- function (dt, weekday) {
            # Mean
          dt_Date <- dt %>% group_by(Location.ID, date, hour) %>% summarize(nWazeJam = sum(nWazeJam),
                                                                         Obs = sum(Obs),
                                                                         nWazeAccident = sum(nWazeAccident),
                                                                         nHazardOnShoulder = sum(nHazardOnShoulder),
                                                                         nHazardOnRoad = sum(nHazardOnRoad),
                                                                         nHazardWeather = sum(nHazardWeather)
          ) %>% mutate(DayofWeek = weekdays(date), Date = paste(date, DayofWeek), Month = month(date)) %>%  left_join(unique(dt[, c("date","EventType","StartTime","EndTime")]), by = "date") %>% mutate(background_col = ifelse(EventType == "NoEvent", NA,ifelse(EventType == "Football","#E69F00",ifelse(EventType == "Concert", "#56B4E9", "chartreuse4" ))))
          
          dt_Date$Month <- factor(dt_Date$Month) 
          levels(dt_Date$Month) <- list("April" = 4, "May" = 5, "June" = 6, "July" = 7, "August" = 8, "September" = 9)
          
          # Military time to regular HH:MM, and then to decimals.
          dt_Date$StartTime <- hm_to_num(mil_to_hm(dt_Date$StartTime))
          dt_Date$EndTime <- hm_to_num(mil_to_hm(dt_Date$EndTime))
          
          dt_visual <-  dt_Date %>% filter(DayofWeek == weekday)
          
          ggplot(dt_visual, aes(x = hour, y = Obs, color = EventType, group = date)) + 
            # geom_point(alpha = 0.5, color = "red") +
            geom_line(alpha = 0.5) +
            geom_rect(data = dt_visual,aes(xmin = ave(dt_visual$StartTime, dt_visual$date, FUN = function(x) mean(x)),xmax = ave(dt_visual$EndTime, dt_visual$date, FUN = function(x) mean(x)), ymin = -Inf,ymax=Inf), fill = dt_visual$background_col, alpha= 0.2/100, inherit.aes = FALSE) +
            facet_wrap(~ Date) +
            ylab("Total EDT Accident") +
            xlab("Hour of Day") +
            scale_color_manual(name = NULL, values=c("NoEvent" = "#999999", "Football" = "#E69F00", "Concert" = "#56B4E9", "Soccer" = "chartreuse4")) +
            ggtitle(paste("Fedex Field,", buf,"Mile Buffer,", weekday)) 
}


weekday = "Sunday"
plotSumJam(dt,weekday)
plotSumCrash(dt,weekday)
plotSumEDT(dt,weekday)

# weekday = "Saturday"
# plotSumJam(dt,weekday)
# 
# weekday = "Tuesday"
# plotSumJam(dt,weekday)
# 
# weekday = "Wednesday"
# plotSumJam(dt,weekday)
```

### Hourly By Day of Week - 1 mile buffer

```{r, fig.width=12, fig.height=12}
buf = 1
dt <- GridDataSE(loc,buf,daterange,col.names)

weekday = "Sunday"
plotSumJam(dt,weekday)
plotSumCrash(dt,weekday)
plotSumEDT(dt,weekday)

# weekday = "Saturday"
# plotAvgJam(dt,weekday)
# 
# weekday = "Tuesday"
# plotAvgJam(dt,weekday)
# 
# weekday = "Wednesday"
# plotAvgJam(dt,weekday)
```

### Hourly By Day of Week - half mile buffer

```{r, fig.width=12, fig.height=12}
buf = 0.5
dt <- GridDataSE(loc,buf,daterange,col.names)

weekday = "Sunday"
plotSumJam(dt,weekday)
plotSumCrash(dt,weekday)
plotSumEDT(dt,weekday)

# weekday = "Saturday"
# plotAvgJam(dt,weekday)
# 
# weekday = "Tuesday"
# plotAvgJam(dt,weekday)
# 
# weekday = "Wednesday"
# plotAvgJam(dt,weekday)
```

## Comparing all Matrics

```{r eval = F}
dt_EventType <- dt %>% group_by(EventType, hour) %>% summarize(nWazeJam = mean(nWazeJam),
                                                         Obs = mean(Obs),
                                                         nWazeAccident = mean(nWazeAccident),
                                                         nHazardOnShoulder = mean(nHazardOnShoulder),
                                                         nHazardOnRoad = mean(nHazardOnRoad),
                                                         nHazardWeather = mean(nHazardWeather)
                                                         )

ggplot(dt_EventType, aes(x = hour, y = Obs)) + geom_point() + geom_line() + facet_wrap(~ EventType) + ylab("Average EDT Accident") + ggtitle(paste(loc, buf,"mile buffer"))

ggplot(dt_EventType, aes(x = hour, y = nWazeAccident)) + geom_point() + geom_line() + facet_wrap(~ EventType) + ylab("Average Waze Accident") + ggtitle(paste(loc, buf,"mile buffer"))

ggplot(dt_EventType, aes(x = hour, y = nWazeJam)) + geom_point() + geom_line() + facet_wrap(~ EventType) + ylab("Average Waze Jam") + ggtitle(paste(loc, buf,"mile buffer"))

ggplot(dt_EventType, aes(x = hour, y = nHazardOnShoulder)) + geom_point() + geom_line() + facet_wrap(~ EventType) + ylab("Average Waze Hazard On Shoulder") + ggtitle(paste(loc, buf,"mile buffer"))

ggplot(dt_EventType, aes(x = hour, y = nHazardOnRoad)) + geom_point() + geom_line() + facet_wrap(~ EventType) + ylab("Average Waze Hazard On Road") + ggtitle(paste(loc, buf,"mile buffer"))
```