---
title: "Special Events Visualization - Part 1"
author: "Ci (Jessie) Yang"
date: "`r paste('Last Updated:',format(Sys.time(), '%B %d, %Y'))`"
output:
  html_document: default
pdf_document:
  fig_caption: yes
fig_height: 5
fig_width: 8
---

This is an example of visuals that we can create visualizations for a specific location and buffer zone around the location where we have model outputs on.

```{r setup, include=FALSE}
#set options
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE)
options(scipen = 20)

source2 <- function(file, start, end, ...) {
    file.lines <- scan(file, what=character(), skip=start-1, nlines=end-start+1, sep='\n')
    file.lines.collapsed <- paste(file.lines, collapse='\n')
    source(textConnection(file.lines.collapsed), ...)
}

source2("SpecialEventsAnalysis.R", 3, 79) # setting
source2("SpecialEventsAnalysis.R", 233, 272) # Data process

# Plotting functions.
meanfun <- function(dt){
  dt_Date <- dt %>% group_by(Location.ID, date, hour) %>% 
            summarize(nWazeJam = mean(nWazeJam),
                      Obs = mean(Obs),
                      nWazeAccident = mean(nWazeAccident),
                      nHazardOnShoulder = mean(nHazardOnShoulder),
                      nHazardOnRoad = mean(nHazardOnRoad),
                      nHazardWeather = mean(nHazardWeather),
                      nEDTFatal = mean(nEDTFatal),
                      nEDTCrashWPedBikeInv = mean(nEDTCrashWPedBikeInv),
                      nEDTCrashWSchoolBusInv = mean(nEDTCrashWSchoolBusInv),
                      nEDTVehCount = mean(nEDTVehCount)
                      )
  dt_Date
}

sumfun <- function(dt){
  dt_Date <- dt %>% group_by(Location.ID, date, hour) %>% 
            summarize(nWazeJam = sum(nWazeJam),
                      Obs = sum(Obs),
                      nWazeAccident = sum(nWazeAccident),
                      nHazardOnShoulder = sum(nHazardOnShoulder),
                      nHazardOnRoad = sum(nHazardOnRoad),
                      nHazardWeather = sum(nHazardWeather),
                      nEDTFatal = sum(nEDTFatal),
                      nEDTCrashWPedBikeInv = sum(nEDTCrashWPedBikeInv),
                      nEDTCrashWSchoolBusInv = sum(nEDTCrashWSchoolBusInv),
                      nEDTVehCount = sum(nEDTVehCount)
                      )
  dt_Date
}

maxfun <- function(dt){
  dt_Date <- dt %>% group_by(Location.ID, date, hour) %>% 
            summarize(nWazeJam = max(nWazeJam),
                      Obs = max(Obs),
                      nWazeAccident = max(nWazeAccident),
                      nHazardOnShoulder = max(nHazardOnShoulder),
                      nHazardOnRoad = max(nHazardOnRoad),
                      nHazardWeather = max(nHazardWeather),
                      nEDTFatal = max(nEDTFatal),
                      nEDTCrashWPedBikeInv = max(nEDTCrashWPedBikeInv),
                      nEDTCrashWSchoolBusInv = max(nEDTCrashWSchoolBusInv),
                      nEDTVehCount = max(nEDTVehCount)
                      )
  dt_Date
}

plotMonth <- function(dt, loc, weekday, statistic, metrics){
        
        location <- unique(SpecialEvents$Location[SpecialEvents$Location.ID == loc])
        dt_Date <- if (statistic == "Total") sumfun(dt) else if (statistic == "Average") meanfun(dt) else maxfun(dt)
            
          # Process data to include more information of the date and event
         dt_Date <- dt_Date %>% mutate(DayofWeek = weekdays(date), Date = paste(date, DayofWeek), Month = month(date)) %>%  left_join(unique(dt[, c("date","EventType","StartTime","EndTime")]), by = "date") %>% mutate(background_col = ifelse(EventType == "NoEvent", NA,ifelse(EventType == "Football","#E69F00",ifelse(EventType == "Concert", "#56B4E9", ifelse(EventType == "Soccer", "chartreuse4","red")))))
          
          dt_Date$Month <- factor(dt_Date$Month) 
          levels(dt_Date$Month) <- list("April" = 4, "May" = 5, "June" = 6, "July" = 7, "August" = 8, "September" = 9)
          
          # Military time to regular HH:MM, and then to decimals.
          dt_Date$StartTime <- hm_to_num(mil_to_hm(dt_Date$StartTime))
          dt_Date$EndTime <- hm_to_num(mil_to_hm(dt_Date$EndTime))
          
          dt_visual <-  dt_Date %>% filter(DayofWeek == weekday)
          
          # Choose data
          p <- if(metrics == "EDT Accident") ggplot(dt_visual, aes(x = hour, y = Obs, color = EventType, group = date)) else if (metrics == "Waze Jam") ggplot(dt_visual, aes(x = hour, y = nWazeJam, color = EventType, group = date)) else if (metrics == "Waze Accident") ggplot(dt_visual, aes(x = hour, y = nWazeAccident, color = EventType, group = date))  else if (metrics == "Waze Hazard On Shoulder") ggplot(dt_visual, aes(x = hour, y = nHazardOnShoulder, color = EventType, group = date)) else if (metrics == "Waze Hazard On Road") ggplot(dt_visual, aes(x = hour, y = nHazardOnRoad, color = EventType, group = date)) else if (metrics == "EDT Fatals") ggplot(dt_visual, aes(x = hour, y = nEDTFatal, color = EventType, group = date)) else if (metrics == "EDT Crashes with Ped or Bike Involvement") ggplot(dt_visual, aes(x = hour, y = nEDTCrashWPedBikeInv, color = EventType, group = date))  else if (metrics == "EDT Crashes with School Bus Involvement") ggplot(dt_visual, aes(x = hour, y = nEDTCrashWSchoolBusInv, color = EventType, group = date)) else  ggplot(dt_visual, aes(x = hour, y = nEDTVehCount, color = EventType, group = date))

          p + 
            # geom_point(alpha = 0.5, color = "red") +
            geom_line(alpha = 0.5) +
            geom_rect(data = dt_visual,aes(xmin = ave(dt_visual$StartTime, dt_visual$date, FUN = function(x) mean(x)),xmax = ave(dt_visual$EndTime, dt_visual$date, FUN = function(x) mean(x)), ymin = -Inf,ymax=Inf), fill = dt_visual$background_col, alpha= 0.2/100, inherit.aes = FALSE) +
            facet_wrap(~ Month) +
            xlab("Hour of Day") +
            ylab(paste(statistic, metrics)) +
            scale_color_manual(name = NULL, values=c("NoEvent" = "#999999", "Football" = "#E69F00", "Concert" = "#56B4E9", "Soccer" = "chartreuse4", "Baseball" = "red")) +
            ggtitle(paste(location,",", buf,"Mile Buffer,", weekday)) 
}

plotWeekdayHeatMap <- function(dt, loc, metrics){
  
        location <- unique(SpecialEvents$Location[SpecialEvents$Location.ID == loc])
        
        dt_Date <- dt %>% group_by(Location.ID, DayofWeek, EventType, hour) %>% 
            summarize(numDay = length(unique(date)),
                      nWazeJam = ifelse(sum(nWazeJam) == 0, NA, sum(nWazeJam)/numDay),
                      Obs = ifelse(sum(Obs) == 0, NA, sum(Obs)/numDay),
                      nWazeAccident = ifelse(sum(nWazeAccident) == 0, NA,sum(nWazeAccident)/numDay),
                      nHazardOnShoulder = ifelse(sum(nHazardOnShoulder) == 0, NA,sum(nHazardOnShoulder)/numDay),
                      nHazardOnRoad = ifelse(sum(nHazardOnRoad) == 0, NA,sum(nHazardOnRoad)/numDay),
                      nHazardWeather = ifelse(sum(nHazardWeather) == 0, NA,sum(nHazardWeather)/numDay),
                      nEDTFatal = ifelse(sum(nEDTFatal) == 0, NA,sum(nEDTFatal)/numDay),
                      nEDTCrashWPedBikeInv = ifelse(sum(nEDTCrashWPedBikeInv) == 0, NA,sum(nEDTCrashWPedBikeInv)/numDay),
                      nEDTCrashWSchoolBusInv = ifelse(sum(nEDTCrashWSchoolBusInv) == 0, NA,sum(nEDTCrashWSchoolBusInv)/numDay),
                      nEDTVehCount = ifelse(sum(nEDTVehCount) == 0, NA,sum(nEDTVehCount)/numDay)
                      )
          
          dt_Date$DayofWeek <- factor(dt_Date$DayofWeek) 
          levels(dt_Date$DayofWeek) <- list("Monday" = "Monday","Tuesday" = "Tuesday","Wednesday" = "Wednesday","Thursday" = "Thursday", "Friday" = "Friday", "Saturday" = "Saturday", "Sunday" = "Sunday")
          
          dt_visual <-  dt_Date 
          # %>% filter(DayofWeek == weekday)
         
          # Heatmap
          p <- if (metrics == "EDT Accident") {
            ggplot(dt_visual, aes(x = hour, y = EventType, fill = Obs)) 
            # + geom_text(aes(label = round(ifelse(Obs > 0, Obs, NA)))) 
          } else if (metrics == "Waze Jam") {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nWazeJam)) 
            # + geom_text(aes(label = round(ifelse(nWazeJam > 0, nWazeJam, NA)))) 
          } else if (metrics == "Waze Accident") {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nWazeAccident)) 
            # + geom_text(aes(label = round(ifelse(nWazeAccident > 0, nWazeAccident, NA))))
          } else if (metrics == "Waze Hazard On Shoulder") {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nHazardOnShoulder)) 
            # + geom_text(aes(label = round(ifelse(nHazardOnShoulder > 0, nHazardOnShoulder, NA)))) 
          } else if (metrics == "Waze Hazard On Road") {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nHazardOnRoad)) 
            # +  geom_text(aes(label = round(ifelse(nHazardOnRoad > 0, nHazardOnRoad, NA))))
          } else if (metrics == "EDT Fatals") {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nEDTFatal)) 
            # + geom_text(aes(label = round(ifelse(nEDTFatal > 0, nEDTFatal, NA))))
          } else if (metrics == "EDT Crashes with Ped or Bike Involvement") {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nEDTCrashWPedBikeInv)) 
            # + geom_text(aes(label = round(ifelse(nEDTCrashWPedBikeInv > 0, nEDTCrashWPedBikeInv, NA))))
          } else if (metrics == "EDT Crashes with School Bus Involvement") {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nEDTCrashWSchoolBusInv)) 
            # + geom_text(aes(label = round(ifelse(nEDTCrashWSchoolBusInv > 0, nEDTCrashWSchoolBusInv, NA))))
         } else  {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nEDTVehCount)) 
           # + geom_text(aes(label = round(ifelse(nEDTVehCount > 0, nEDTVehCount, NA))))
           }
         p + geom_tile(color = "white", aes(height = 0.5)) + # background colours are mapped according to the value column
            scale_fill_gradient(low = "#fee08b", high = "#d53e4f", na.value = "white", space = "Lab", guide = "colourbar") + # determine the colour
            facet_wrap(~ DayofWeek) +
            ggtitle(paste0(location,", ",buf, " Mile Buffer")) + xlab("Hour") + ylab("") +
            theme(legend.title=element_text(face="bold", size=14)) + 
            labs(fill= paste0(metrics))
}

plotWeekdayendHeatMap <- function(dt, loc, metrics){
  
        location <- unique(SpecialEvents$Location[SpecialEvents$Location.ID == loc]
        
        dt$WeekDayEnd <- ifelse(dt$DayofWeek %in% c("Monday" ,"Tuesday" ,"Wednesday" ,"Thursday" , "Friday"), "Weekday","Weekend")
          
        dt_Date <- dt %>% group_by(Location.ID, WeekDayEnd, EventType, hour) %>% 
            summarize(numDay = length(unique(date)),
                      nWazeJam = ifelse(sum(nWazeJam) == 0, NA, sum(nWazeJam)/numDay),
                      Obs = ifelse(sum(Obs) == 0, NA, sum(Obs)/numDay),
                      nWazeAccident = ifelse(sum(nWazeAccident) == 0, NA,sum(nWazeAccident)/numDay),
                      nHazardOnShoulder = ifelse(sum(nHazardOnShoulder) == 0, NA,sum(nHazardOnShoulder)/numDay),
                      nHazardOnRoad = ifelse(sum(nHazardOnRoad) == 0, NA,sum(nHazardOnRoad)/numDay),
                      nHazardWeather = ifelse(sum(nHazardWeather) == 0, NA,sum(nHazardWeather)/numDay),
                      nEDTFatal = ifelse(sum(nEDTFatal) == 0, NA,sum(nEDTFatal)/numDay),
                      nEDTCrashWPedBikeInv = ifelse(sum(nEDTCrashWPedBikeInv) == 0, NA,sum(nEDTCrashWPedBikeInv)/numDay),
                      nEDTCrashWSchoolBusInv = ifelse(sum(nEDTCrashWSchoolBusInv) == 0, NA,sum(nEDTCrashWSchoolBusInv)/numDay),
                      nEDTVehCount = ifelse(sum(nEDTVehCount) == 0, NA,sum(nEDTVehCount)/numDay)
                      )
          
          dt_visual <-  dt_Date 
          # %>% filter(DayofWeek == weekday)
         
          # Heatmap
          # Heatmap
          p <- if (metrics == "EDT Accident") {
            ggplot(dt_visual, aes(x = hour, y = EventType, fill = Obs)) 
            # + geom_text(aes(label = round(ifelse(Obs > 0, Obs, NA)))) 
          } else if (metrics == "Waze Jam") {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nWazeJam)) 
            # + geom_text(aes(label = round(ifelse(nWazeJam > 0, nWazeJam, NA)))) 
          } else if (metrics == "Waze Accident") {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nWazeAccident)) 
            # + geom_text(aes(label = round(ifelse(nWazeAccident > 0, nWazeAccident, NA))))
          } else if (metrics == "Waze Hazard On Shoulder") {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nHazardOnShoulder)) 
            # + geom_text(aes(label = round(ifelse(nHazardOnShoulder > 0, nHazardOnShoulder, NA)))) 
          } else if (metrics == "Waze Hazard On Road") {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nHazardOnRoad)) 
            # +  geom_text(aes(label = round(ifelse(nHazardOnRoad > 0, nHazardOnRoad, NA))))
          } else if (metrics == "EDT Fatals") {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nEDTFatal)) 
            # + geom_text(aes(label = round(ifelse(nEDTFatal > 0, nEDTFatal, NA))))
          } else if (metrics == "EDT Crashes with Ped or Bike Involvement") {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nEDTCrashWPedBikeInv)) 
            # + geom_text(aes(label = round(ifelse(nEDTCrashWPedBikeInv > 0, nEDTCrashWPedBikeInv, NA))))
          } else if (metrics == "EDT Crashes with School Bus Involvement") {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nEDTCrashWSchoolBusInv)) 
            # + geom_text(aes(label = round(ifelse(nEDTCrashWSchoolBusInv > 0, nEDTCrashWSchoolBusInv, NA))))
         } else  {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nEDTVehCount)) 
           # + geom_text(aes(label = round(ifelse(nEDTVehCount > 0, nEDTVehCount, NA))))
           }
         p + geom_tile(color = "white", aes(height = 0.5)) + # background colours are mapped according to the value column
            scale_fill_gradient(low = "#fee08b", high = "#d53e4f", na.value = "white", space = "Lab", guide = "colourbar") + # determine the colour
            facet_wrap(~ WeekDayEnd, ncol= 1) +
            ggtitle(paste0(location,", ",buf, " Mile Buffer")) + xlab("Hour") + ylab("") +
            theme(legend.title=element_text(face="bold", size=14)) + 
            labs(fill= paste0(metrics))
}



plotDateHeatMap <- function(dt, loc, weekday, metrics){
  
        location <- unique(SpecialEvents$Location[SpecialEvents$Location.ID == loc])
        
        dt_Date <- dt %>% group_by(Location.ID, DayofWeek, EventType, hour) %>% 
            summarize(numDay = length(unique(date)),
                      nWazeJam = ifelse(sum(nWazeJam) == 0, NA, sum(nWazeJam)/numDay),
                      Obs = ifelse(sum(Obs) == 0, NA, sum(Obs)/numDay),
                      nWazeAccident = ifelse(sum(nWazeAccident) == 0, NA,sum(nWazeAccident)/numDay),
                      nHazardOnShoulder = ifelse(sum(nHazardOnShoulder) == 0, NA,sum(nHazardOnShoulder)/numDay),
                      nHazardOnRoad = ifelse(sum(nHazardOnRoad) == 0, NA,sum(nHazardOnRoad)/numDay),
                      nHazardWeather = ifelse(sum(nHazardWeather) == 0, NA,sum(nHazardWeather)/numDay),
                      nEDTFatal = ifelse(sum(nEDTFatal) == 0, NA,sum(nEDTFatal)/numDay),
                      nEDTCrashWPedBikeInv = ifelse(sum(nEDTCrashWPedBikeInv) == 0, NA,sum(nEDTCrashWPedBikeInv)/numDay),
                      nEDTCrashWSchoolBusInv = ifelse(sum(nEDTCrashWSchoolBusInv) == 0, NA,sum(nEDTCrashWSchoolBusInv)/numDay),
                      nEDTVehCount = ifelse(sum(nEDTVehCount) == 0, NA,sum(nEDTVehCount)/numDay)
                      )
        
          dt_visual <-  dt_Date  %>% filter(DayofWeek == weekday)
         
          # Heatmap
          p <- if (metrics == "EDT Accident") {
            ggplot(dt_visual, aes(x = hour, y = EventType, fill = Obs)) 
            # + geom_text(aes(label = round(ifelse(Obs > 0, Obs, NA)))) 
          } else if (metrics == "Waze Jam") {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nWazeJam)) 
            # + geom_text(aes(label = round(ifelse(nWazeJam > 0, nWazeJam, NA)))) 
          } else if (metrics == "Waze Accident") {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nWazeAccident)) 
            # + geom_text(aes(label = round(ifelse(nWazeAccident > 0, nWazeAccident, NA))))
          } else if (metrics == "Waze Hazard On Shoulder") {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nHazardOnShoulder)) 
            # + geom_text(aes(label = round(ifelse(nHazardOnShoulder > 0, nHazardOnShoulder, NA)))) 
          } else if (metrics == "Waze Hazard On Road") {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nHazardOnRoad)) 
            # +  geom_text(aes(label = round(ifelse(nHazardOnRoad > 0, nHazardOnRoad, NA))))
          } else if (metrics == "EDT Fatals") {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nEDTFatal)) 
            # + geom_text(aes(label = round(ifelse(nEDTFatal > 0, nEDTFatal, NA))))
          } else if (metrics == "EDT Crashes with Ped or Bike Involvement") {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nEDTCrashWPedBikeInv)) 
            # + geom_text(aes(label = round(ifelse(nEDTCrashWPedBikeInv > 0, nEDTCrashWPedBikeInv, NA))))
          } else if (metrics == "EDT Crashes with School Bus Involvement") {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nEDTCrashWSchoolBusInv)) 
            # + geom_text(aes(label = round(ifelse(nEDTCrashWSchoolBusInv > 0, nEDTCrashWSchoolBusInv, NA))))
         } else  {ggplot(dt_visual, aes(x = hour, y = EventType, fill = nEDTVehCount)) 
           # + geom_text(aes(label = round(ifelse(nEDTVehCount > 0, nEDTVehCount, NA))))
           }

         p + geom_tile(color = "white", aes(height = 0.5)) + # background colours are mapped according to the value column
            scale_fill_gradient(low = "#fee08b", high = "#d53e4f", na.value = "white", space = "Lab", guide = "colourbar") + # determine the colour
            ggtitle(paste0(location,", ",buf, " Mile Buffer, ", weekday)) + xlab("Hour") + ylab("") +
            theme(legend.title=element_text(face="bold", size=14)) + 
            labs(fill= paste0(metrics))
}


plotDate <- function(dt, loc, weekday, statistic, metrics){
        
        location <- unique(SpecialEvents$Location[SpecialEvents$Location.ID == loc])
        dt_Date <- if (statistic == "Total") sumfun(dt) else if (statistic == "Average") meanfun(dt) else maxfun(dt)
            
          # Process data to include more information of the date and event
         dt_Date <- dt_Date %>% mutate(DayofWeek = weekdays(date), Date = paste(date, DayofWeek), Month = month(date)) %>%  left_join(unique(dt[, c("date","EventType","StartTime","EndTime")]), by = "date") %>% mutate(background_col = ifelse(EventType == "NoEvent", NA,ifelse(EventType == "Football","#E69F00",ifelse(EventType == "Concert", "#56B4E9", ifelse(EventType == "Soccer", "chartreuse4","red")))))
          
          dt_Date$Month <- factor(dt_Date$Month) 
          levels(dt_Date$Month) <- list("April" = 4, "May" = 5, "June" = 6, "July" = 7, "August" = 8, "September" = 9)
          
          # Military time to regular HH:MM, and then to decimals.
          dt_Date$StartTime <- hm_to_num(mil_to_hm(dt_Date$StartTime))
          dt_Date$EndTime <- hm_to_num(mil_to_hm(dt_Date$EndTime))
          
          dt_visual <-  dt_Date %>% filter(DayofWeek == weekday)
          
          # Choose data
          p <- if(metrics == "EDT Accident") ggplot(dt_visual, aes(x = hour, y = Obs, color = EventType, group = date)) else if (metrics == "Waze Jam") ggplot(dt_visual, aes(x = hour, y = nWazeJam, color = EventType, group = date)) else if (metrics == "Waze Accident") ggplot(dt_visual, aes(x = hour, y = nWazeAccident, color = EventType, group = date))  else if (metrics == "Waze Hazard On Shoulder") ggplot(dt_visual, aes(x = hour, y = nHazardOnShoulder, color = EventType, group = date)) else if (metrics == "Waze Hazard On Road") ggplot(dt_visual, aes(x = hour, y = nHazardOnRoad, color = EventType, group = date)) else if (metrics == "EDT Fatals") ggplot(dt_visual, aes(x = hour, y = nEDTFatal, color = EventType, group = date)) else if (metrics == "EDT Crashes with Ped or Bike Involvement") ggplot(dt_visual, aes(x = hour, y = nEDTCrashWPedBikeInv, color = EventType, group = date))  else if (metrics == "EDT Crashes with School Bus Involvement") ggplot(dt_visual, aes(x = hour, y = nEDTCrashWSchoolBusInv, color = EventType, group = date)) else  ggplot(dt_visual, aes(x = hour, y = nEDTVehCount, color = EventType, group = date))

          p + 
            # geom_point(alpha = 0.5, color = "red") +
            geom_line(alpha = 0.5) +
            geom_rect(data = dt_visual,aes(xmin = ave(dt_visual$StartTime, dt_visual$date, FUN = function(x) mean(x)),xmax = ave(dt_visual$EndTime, dt_visual$date, FUN = function(x) mean(x)), ymin = -Inf,ymax=Inf), fill = dt_visual$background_col, alpha= 0.2/100, inherit.aes = FALSE) +
            facet_wrap(~ Date, ncol= 1) +
            xlab("Hour of Day") +
            ylab(paste(statistic, metrics)) +
            scale_color_manual(name = NULL, values=c("NoEvent" = "#999999", "Football" = "#E69F00", "Concert" = "#56B4E9", "Soccer" = "chartreuse4", "Baseball" = "red")) +
            ggtitle(paste(location,",", buf,"Mile Buffer,", weekday)) 
}

plotDay <- function(dt, loc, days, statistic, metrics){
        
        location <- unique(SpecialEvents$Location[SpecialEvents$Location.ID == loc])
        dt_Date <- if (statistic == "Total") sumfun(dt) else if (statistic == "Average") meanfun(dt) else maxfun(dt)
            
          # Process data to include more information of the date and event
         dt_Date <- dt_Date %>% mutate(DayofWeek = weekdays(date), Date = paste(date, DayofWeek), Month = month(date)) %>%  left_join(unique(dt[, c("date","EventType","StartTime","EndTime")]), by = "date") %>% mutate(background_col = ifelse(EventType == "NoEvent", NA,ifelse(EventType == "Football","#E69F00",ifelse(EventType == "Concert", "#56B4E9", ifelse(EventType == "Soccer", "chartreuse4","red")))))
          
          dt_Date$Month <- factor(dt_Date$Month) 
          levels(dt_Date$Month) <- list("April" = 4, "May" = 5, "June" = 6, "July" = 7, "August" = 8, "September" = 9)
          
          # Military time to regular HH:MM, and then to decimals.
          dt_Date$StartTime <- hm_to_num(mil_to_hm(dt_Date$StartTime))
          dt_Date$EndTime <- hm_to_num(mil_to_hm(dt_Date$EndTime))
          
          dt_visual <-  dt_Date %>% filter(date %in% as.Date(days))
          
          # Choose data
          p <- if(metrics == "EDT Accident") ggplot(dt_visual, aes(x = hour, y = Obs, color = EventType, group = date)) else if (metrics == "Waze Jam") ggplot(dt_visual, aes(x = hour, y = nWazeJam, color = EventType, group = date)) else if (metrics == "Waze Accident") ggplot(dt_visual, aes(x = hour, y = nWazeAccident, color = EventType, group = date))  else if (metrics == "Waze Hazard On Shoulder") ggplot(dt_visual, aes(x = hour, y = nHazardOnShoulder, color = EventType, group = date)) else if (metrics == "Waze Hazard On Road") ggplot(dt_visual, aes(x = hour, y = nHazardOnRoad, color = EventType, group = date)) else if (metrics == "EDT Fatals") ggplot(dt_visual, aes(x = hour, y = nEDTFatal, color = EventType, group = date)) else if (metrics == "EDT Crashes with Ped or Bike Involvement") ggplot(dt_visual, aes(x = hour, y = nEDTCrashWPedBikeInv, color = EventType, group = date))  else if (metrics == "EDT Crashes with School Bus Involvement") ggplot(dt_visual, aes(x = hour, y = nEDTCrashWSchoolBusInv, color = EventType, group = date)) else  ggplot(dt_visual, aes(x = hour, y = nEDTVehCount, color = EventType, group = date))

          p + 
            # geom_point(alpha = 0.5, color = "red") +
            geom_line(alpha = 0.5) +
            geom_rect(data = dt_visual,aes(xmin = ave(dt_visual$StartTime, dt_visual$date, FUN = function(x) mean(x)),xmax = ave(dt_visual$EndTime, dt_visual$date, FUN = function(x) mean(x)), ymin = -Inf,ymax=Inf), fill = dt_visual$background_col, alpha= 0.2/100, inherit.aes = FALSE) +
            facet_wrap(~ date, ncol = 1) +
            xlab("Hour of Day") +
            ylab(paste(statistic, metrics)) +
            scale_color_manual(name = NULL, values=c("NoEvent" = "#999999", "Football" = "#E69F00", "Concert" = "#56B4E9", "Soccer" = "chartreuse4", "Baseball" = "red")) +
            ggtitle(paste(location,",", buf,"Mile Buffer")) 
          
}


```

# Time Series Plots

We used the AllModel30.csv output for this analysis.

Metrics we use include:

* Number of EDT Accidents (Model Dependent Variable)
* Number of Waze Accidents
* Number of Waze Jam
* Number of Hazard On Shoulder (nHazardOnShoulder)
* Number of Hazard On Road (nHazardOnRoad)
* Number of Hazard Weather (nHazardWeather) - is it # of hazard or weather? According to Michelle, this should be one of the three categories of Waze event (Accidents, Jam, and Hazard/Weather). However, in the dataset, nHazardWeather is smaller than sum of the sub-categories (e.g., nHazardOnShoulder, nHazardOnRoad). This needs further investigation.
* Probability of NonCrash and Crash - real prob from RF model (needs further investigation/decision making)
The probability has converted to based on the threshold ~ 0.2 using the built-in function in RF.

To do 8/23/2018:

* total/average/max (point/line) number of accidents/jams for every day (done)
    * plot on top of each other for the NoEvents with a different color. Use other colors for a type of special events. 4/23 (4/16 easter), 5/21 (5/29 memorial day, 5/14 mother's day), 7/02 weekends near holidays. 
    * Do both 3-miles and 1-mile. Removing zeros. 
    * Panels by month.
    * add shadebox for event duration
* add function for visuals (done)
* number of accidents per gridcell per hour (done)
* boxplot of accidents over 42 grid cells. (done)
* look at hour aggregation
* heatmap for each hour
* smaller buffer zone, 0.25, 0.5, 1 mile radius 
    + update SpecialEventsExpand dataset (done)
    + update the analysis and visuals (done)
* Easten shore beach - Bay Bridge area (7/4, 9/4 Labor Day, 3lanes both ways) for a typical holiday pattern
* follow up with Erika tomorrow (done)
* write one plot function in stead of multiple plot functions. (done)

Dave's feedback 8/29/2018:

* try large buffer, e.g., 5 miles (Dan's suggestion: work with individual lag/lons instead of grid cells)
* broader patterns - can we add baseball game,(Michelle's comments: baseball games are usually hosted in urban areas with intercity activities and good public transportation, less parking. Erika - How far aways is the parking. Look that the parking in the buffer).
* Two days of week (Saturday, Wednesday)
* Extend the 23 hours through another 6 hours, change time window from 5 am to 5 am.
* Other types of metric, Hazards on the roads, car stopped on shoulders. Car stopped on shoulders.
* Other location: for exmaple, University football games.
* Can we try to make this a little more of hot spots, can we do this on a grid cell level.

## HeatMap

### 5 mile buffer zone
```{r eval = T}
loc = "SE1"
buf = 5
weekday = "Sunday"
dt <- GridDataSE(loc,buf,alldays,col.names)
plotDateHeatMap(dt, loc, weekday, "EDT Accident") # average values
plotDateHeatMap(dt, loc, weekday, "Waze Jam")
plotDateHeatMap(dt, loc, weekday, "Waze Accident")

plotWeekdayHeatMap(dt, loc, "Waze Jam")
plotWeekdayHeatMap(dt, loc, "EDT Accident")
plotWeekdayHeatMap(dt, loc, "Waze Accident")

loc = "SE2"
buf = 5
weekday = "Sunday"
plotDateHeatMap(dt, loc, weekday, "EDT Accident")
plotDateHeatMap(dt, loc, weekday, "Waze Jam")

# plotWeekdayHeatMap(dt, loc, "Waze Jam")
# plotWeekdayHeatMap(dt, loc, "EDT Accident")
# plotWeekdayHeatMap(dt, loc, "Waze Accident")

plotWeekdayendHeatMap(dt, loc, "Waze Jam")
plotWeekdayendHeatMap(dt, loc, "EDT Accident")
plotWeekdayendHeatMap(dt, loc, "Waze Accident")

# make the bars shorter
# weekday/weekend classification - baseball only.
# Fedex field: football
```

## Example Days in Presentation

### 5 mile buffer zone - Total
```{r eval = T}
loc = "SE1"
buf = 5 # both 3 and 5 miles buffer
dt <- GridDataSE(loc,buf,alldays,col.names)
days = c("2017-09-10","2017-09-17")
statistic = "Total" #c("Total","Average","Maximum")
metrics = "EDT Accident" #c("EDT Accident","Waze Accident","Waze Jam","Waze Hazard On Shoulder","Waze Hazard On Road","EDT Fatals","EDT Crashes with Ped or Bike Involvement","EDT Crashes with School Bus Involvement","EDT Vehicle Counts")

plotDay(dt, loc, days, statistic, "Waze Accident")
plotDay(dt, loc, days, statistic, "Waze Jam")
plotDay(dt, loc, days, statistic, "Waze Hazard On Shoulder")
plotDay(dt, loc, days, statistic, "Waze Hazard On Road")
plotDay(dt, loc, days, statistic, "EDT Accident")
plotDay(dt, loc, days, statistic, "EDT Fatals")
plotDay(dt, loc, days, statistic, "EDT Crashes with Ped or Bike Involvement")
plotDay(dt, loc, days, statistic, "EDT Crashes with School Bus Involvement")
plotDay(dt, loc, days, statistic, "EDT Vehicle Counts")

# stackle plots in another orientation, on top of each other.

```


## Hourly By Event Type

### 5 mile buffer zone - Total
```{r eval = T}
loc = "SE2"
buf = 5
dt <- GridDataSE(loc,buf,alldays,col.names)
weekday = "Saturday"
statistic = "Total" #c("Total","Average","Maximum")
metrics = "EDT Accident" #c("EDT Accident","Waze Accident","Waze Jam","Waze Hazard On Shoulder","Waze Hazard On Road","EDT Fatals","EDT Crashes with Ped or Bike Involvement","EDT Crashes with School Bus Involvement","EDT Vehicle Counts")

plotMonth(dt, loc, weekday, statistic, "Waze Accident")
plotMonth(dt, loc, weekday, statistic, "Waze Jam")
plotMonth(dt, loc, weekday, statistic, "Waze Hazard On Shoulder")
plotMonth(dt, loc, weekday, statistic, "Waze Hazard On Road")
plotMonth(dt, loc, weekday, statistic, "EDT Accident")
plotMonth(dt, loc, weekday, statistic, "EDT Fatals")
plotMonth(dt, loc, weekday, statistic, "EDT Crashes with Ped or Bike Involvement")
plotMonth(dt, loc, weekday, statistic, "EDT Crashes with School Bus Involvement")
plotMonth(dt, loc, weekday, statistic, "EDT Vehicle Counts")

```

### 3 mile buffer zone - Total
```{r eval = T}
loc = "SE2"
buf = 3
dt <- GridDataSE(loc,buf,alldays,col.names)
weekday = "Saturday"
statistic = "Total" #c("Total","Average","Maximum")
metrics = "EDT Accident" #c("EDT Accident","Waze Accident","Waze Jam","Waze Hazard On Shoulder","Waze Hazard On Road")

plotMonth(dt, loc, weekday, statistic, "Waze Accident")
plotMonth(dt, loc, weekday, statistic, "Waze Jam")
plotMonth(dt, loc, weekday, statistic, "Waze Hazard On Shoulder")
plotMonth(dt, loc, weekday, statistic, "Waze Hazard On Road")
plotMonth(dt, loc, weekday, statistic, "EDT Accident")
plotMonth(dt, loc, weekday, statistic, "EDT Fatals")
plotMonth(dt, loc, weekday, statistic, "EDT Crashes with Ped or Bike Involvement")
plotMonth(dt, loc, weekday, statistic, "EDT Crashes with School Bus Involvement")
plotMonth(dt, loc, weekday, statistic, "EDT Vehicle Counts")
```

### 1 mile buffer zone - Total
```{r eval = T}
loc = "SE2"
buf = 1
dt <- GridDataSE(loc,buf,alldays,col.names)
weekday = "Saturday"
statistic = "Total" #c("Total","Average","Maximum")
metrics = "EDT Accident" #c("EDT Accident","Waze Accident","Waze Jam","Waze Hazard On Shoulder","Waze Hazard On Road")

plotMonth(dt, loc, weekday, statistic, "Waze Accident")
plotMonth(dt, loc, weekday, statistic, "Waze Jam")
plotMonth(dt, loc, weekday, statistic, "Waze Hazard On Shoulder")
plotMonth(dt, loc, weekday, statistic, "Waze Hazard On Road")
plotMonth(dt, loc, weekday, statistic, "EDT Accident")
plotMonth(dt, loc, weekday, statistic, "EDT Fatals")
plotMonth(dt, loc, weekday, statistic, "EDT Crashes with Ped or Bike Involvement")
plotMonth(dt, loc, weekday, statistic, "EDT Crashes with School Bus Involvement")
plotMonth(dt, loc, weekday, statistic, "EDT Vehicle Counts")
```


### Hourly By Day of Week - 5 mile buffer

```{r, fig.width=12, fig.height=12}
loc = "SE2"
buf = 5
dt <- GridDataSE(loc,buf,alldays,col.names)
weekday = "Saturday"
statistic = "Total" #c("Total","Average","Maximum")
metrics = "EDT Accident" #c("EDT Accident","Waze Accident","Waze Jam","Waze Hazard On Shoulder","Waze Hazard On Road")

plotDate(dt, loc, weekday, statistic, "Waze Accident")
plotDate(dt, loc, weekday, statistic, "Waze Jam")
plotDate(dt, loc, weekday, statistic, "Waze Hazard On Shoulder")
plotDate(dt, loc, weekday, statistic, "Waze Hazard On Road")
plotDate(dt, loc, weekday, statistic, "EDT Accident")
plotDate(dt, loc, weekday, statistic, "EDT Fatals")
plotDate(dt, loc, weekday, statistic, "EDT Crashes with Ped or Bike Involvement")
plotDate(dt, loc, weekday, statistic, "EDT Crashes with School Bus Involvement")
plotDate(dt, loc, weekday, statistic, "EDT Vehicle Counts")
```
### Hourly By Day of Week - 3 mile buffer

```{r, fig.width=12, fig.height=12}
loc = "SE2"
buf = 3
dt <- GridDataSE(loc,buf,alldays,col.names)
weekday = "Saturday"
statistic = "Total" #c("Total","Average","Maximum")
metrics = "EDT Accident" #c("EDT Accident","Waze Accident","Waze Jam","Waze Hazard On Shoulder","Waze Hazard On Road")

plotDate(dt, loc, weekday, statistic, "Waze Accident")
plotDate(dt, loc, weekday, statistic, "Waze Jam")
plotDate(dt, loc, weekday, statistic, "Waze Hazard On Shoulder")
plotDate(dt, loc, weekday, statistic, "Waze Hazard On Road")
plotDate(dt, loc, weekday, statistic, "EDT Accident")
plotDate(dt, loc, weekday, statistic, "EDT Fatals")
plotDate(dt, loc, weekday, statistic, "EDT Crashes with Ped or Bike Involvement")
plotDate(dt, loc, weekday, statistic, "EDT Crashes with School Bus Involvement")
plotDate(dt, loc, weekday, statistic, "EDT Vehicle Counts")
```

### Hourly By Day of Week - 1 mile buffer

```{r, fig.width=12, fig.height=12}
loc = "SE2"
buf = 1
dt <- GridDataSE(loc,buf,alldays,col.names)
weekday = "Saturday"
statistic = "Total" #c("Total","Average","Maximum")
metrics = "EDT Accident" #c("EDT Accident","Waze Accident","Waze Jam","Waze Hazard On Shoulder","Waze Hazard On Road")

plotDate(dt, loc, weekday, statistic, "Waze Accident")
plotDate(dt, loc, weekday, statistic, "Waze Jam")
plotDate(dt, loc, weekday, statistic, "Waze Hazard On Shoulder")
plotDate(dt, loc, weekday, statistic, "Waze Hazard On Road")
plotDate(dt, loc, weekday, statistic, "EDT Accident")
plotDate(dt, loc, weekday, statistic, "EDT Fatals")
plotDate(dt, loc, weekday, statistic, "EDT Crashes with Ped or Bike Involvement")
plotDate(dt, loc, weekday, statistic, "EDT Crashes with School Bus Involvement")
plotDate(dt, loc, weekday, statistic, "EDT Vehicle Counts")
```

## Comparing all Matrics

```{r eval = F}
dt_EventType <- dt %>% group_by(EventType, hour) %>% summarize(nWazeJam = mean(nWazeJam),
                                                         Obs = mean(Obs),
                                                         nWazeAccident = mean(nWazeAccident),
                                                         nHazardOnShoulder = mean(nHazardOnShoulder),
                                                         nHazardOnRoad = mean(nHazardOnRoad),
                                                         nHazardWeather = mean(nHazardWeather)
                                                         )

ggplot(dt_EventType, aes(x = hour, y = Obs)) + geom_point() + geom_line() + facet_wrap(~ EventType) + ylab("Average EDT Accident") + ggtitle(paste(loc, buf,"mile buffer"))

ggplot(dt_EventType, aes(x = hour, y = nWazeAccident)) + geom_point() + geom_line() + facet_wrap(~ EventType) + ylab("Average Waze Accident") + ggtitle(paste(loc, buf,"mile buffer"))

ggplot(dt_EventType, aes(x = hour, y = nWazeJam)) + geom_point() + geom_line() + facet_wrap(~ EventType) + ylab("Average Waze Jam") + ggtitle(paste(loc, buf,"mile buffer"))

ggplot(dt_EventType, aes(x = hour, y = nHazardOnShoulder)) + geom_point() + geom_line() + facet_wrap(~ EventType) + ylab("Average Waze Hazard On Shoulder") + ggtitle(paste(loc, buf,"mile buffer"))

ggplot(dt_EventType, aes(x = hour, y = nHazardOnRoad)) + geom_point() + geom_line() + facet_wrap(~ EventType) + ylab("Average Waze Hazard On Road") + ggtitle(paste(loc, buf,"mile buffer"))
```


<!-- ### 3 mile buffer zone - Average without zeros -->
<!-- ```{r, eval = F} -->
<!-- buf = 3 -->
<!-- dt <- GridDataSE(loc,buf,alldays,col.names) -->

<!-- # Mean -->
<!-- dt_Date <- dt %>% group_by(Location.ID, date, hour) %>% summarize(nWazeJam = mean(nWazeJam), -->
<!--                                                                Obs = mean(Obs), -->
<!--                                                                nWazeAccident = mean(nWazeAccident), -->
<!--                                                                nHazardOnShoulder = mean(nHazardOnShoulder), -->
<!--                                                                nHazardOnRoad = mean(nHazardOnRoad), -->
<!--                                                                nHazardWeather = mean(nHazardWeather) -->
<!-- ) %>% mutate(DayofWeek = weekdays(date), Date = paste(date, DayofWeek), Month = month(date)) %>%  left_join(unique(dt[, c("date","EventType","StartTime","EndTime")]), by = "date") %>% mutate(background_col = ifelse(EventType == "NoEvent", NA,ifelse(EventType == "Football","#E69F00", "#56B4E9"))) %>% mutate_at(.vars = vars(Obs, nWazeAccident, nWazeJam,nHazardOnShoulder,nHazardOnRoad),.funs = funs(replace(., which(. == 0), NA))) -->

<!-- dt_Date$Month <- factor(dt_Date$Month)  -->
<!-- levels(dt_Date$Month) <- list("April" = 4, "May" = 5, "June" = 6, "July" = 7, "August" = 8, "September" = 9) -->

<!-- # Military time to regular HH:MM, and then to decimals. -->
<!-- dt_Date$StartTime <- hm_to_num(mil_to_hm(dt_Date$StartTime)) -->
<!-- dt_Date$EndTime <- hm_to_num(mil_to_hm(dt_Date$EndTime)) -->

<!-- weekday = "Sunday" -->
<!-- dt_visual <-  dt_Date %>% filter(DayofWeek == weekday) -->

<!-- ggplot(dt_visual, aes(x = hour, y = nWazeJam, color = EventType, group = date)) +  -->
<!--   # geom_point(alpha = 0.5, color = "black") + -->
<!--   geom_line(alpha = 0.5) + -->
<!--   geom_rect(data = dt_visual,aes(xmin = ave(dt_visual$StartTime, dt_visual$date, FUN = function(x) mean(x)),xmax = ave(dt_visual$EndTime, dt_visual$date, FUN = function(x) mean(x)), ymin = -Inf,ymax=Inf), fill = dt_visual$background_col, alpha= 0.2/100, inherit.aes = FALSE) + -->
<!--   facet_wrap(~ Month) + -->
<!--   ylab("Average Waze Jam") + -->
<!--   xlab("Hour of Day") + -->
<!--   scale_color_manual(name = NULL, values=c("NoEvent" = "#999999", "Football" = "#E69F00", "Concert" = "#56B4E9")) + -->
<!--   ggtitle(paste("Fedex Field,", buf,"Mile Buffer,", weekday))  -->
<!-- ``` -->